### 2016-08-02 markdown/draft.RmD
spend far too long trying to sort out things like base directory locations in markdown

---
title: "Decentralisation of Poverty - Replication Analysis"
author: "Jon Minton"
date: "2 August 2016"
output: html_document
---

```{r setup01, include=TRUE, echo = TRUE}
rm(list = ls())
# Code chunk example from http://stackoverflow.com/questions/24585254/working-with-knitr-using-subdirectories
library(knitr)
getwd()
opts_knit$set(root.dir=normalizePath('../'))
```

```{r setup02, include = TRUE, cache = FALSE, echo = TRUE}
getwd()
require(pacman)

pacman::p_load(
  readr, readxl,
  purrr, stringr, tidyr, dplyr,
  rgeos,
  ggplot2, cowplot,
  tmap
)

 source("scripts/do_dz2011_data_management.R")
ls()
centre_distance

```

And now here is a third chunk to check 1) that the working directory has not been reset 
 and 2) to see if the centre_distance object still exists

# Introduction

Here is the contents of centre_distance

## Background

In 2014 Kavanagh, Lee & Pryce produced the AQMeN briefing report [Poverty in Suburbia: Has Glasgow gone the way of American cities?](http://www.research.ed.ac.uk/portal/files/18805757/RB5_poverty_suburbia.pdf). This paper argued that there has been a decentralisation of poverty in Glasgow between 2001 and 2011, and combined a Bayesian spatial statistics approach with the Relative Centralisation Index (RCI) in order to argue firstly that the spatial distribution of poverty has become less decentralised between the two periods, and secondly that the changes in RCE scores are statistically significant.
The briefing paper defined 'poverty' in three ways, each relating to specific UK state benefit receipts for persons of working age:
1. The proportion of local populations receiving Income Support (IS)
2. The proportion of local populations receiving Incapacity Benefit (IB)
3. The proportion of local populations receiving Jobseekers' Allowance (JSA)

These analyses were expanded to the four major Scottish cities - Aberdeen, Dundee, Edinburgh and Glasgow - for a paper to appear in *Annals of the American Association of Geographers*. This paper calculated 95% credible intervals (CrIs) for the difference in RCI from 2001 to 2011 for each of the three measures, and each of the four cities, between 2001 and 2011. It found falls in RCI for each benefit type, and for all four cities, which were statistically significant.

## Purpose

The purposes of this briefing paper are two-fold:
1. To reproduce the analysis in Kavanagh, Lee & Pryce 2016 using the proportion of the population 'income deprived' between 2004 and 2016, using Scottish Index of Multiple Deprivation (SIMD) definitions of income deprivation. This will allow the dependence of Kavanagh, Lee & Pryce's conclusions on the measures and time periods used to be assessed.
2. To use descriptive statistics and data visualisation to better understand the contribution of both changes in the numerators (population counts 'in poverty') and changes in denominators (total population counts) to overall falls in RCI.

# Data and methods

## Data

The SIMD is ...
SIMD data releases for 2004, 2006, 2012 and 2016 were downloaded from the [SIMD website](http://www.gov.scot/Topics/Statistics/SIMD/DataAnalysis). The total population counts, and the population count of the subpopulation defined as 'income deprived', were extracted from each release and combined in a single dataset.

## Methods

Producing the results involved doing the following:
* Defining centroids for each place.
* Identifying the closest centroid for each datazone.
* Reweighting income deprived and total population counts from the SIMD, released at 2001 datazones, onto 2011 datazones.
* etc
* etc

### Calculating nearest centres

The following definitions of city centres were used

Place                 |   2011 Datazone   |  Postcode    |  Description
-------------------   |   -------------   |  ----------- |  -----------
Aberdeen              |   S01006646       |  AB10 1AN    |  Shoe Lane
Glasgow               |   S01010265       |  G1 3BU      |  West End of George Square
Edinburgh             |   S01008677       |  EH1 1BQ     |  31 Waverley Bridge
Dundee                |   S01007705       |  DD1 2AJ     |  Commercial Street
Inverness             |   S01010620       |  IV1 1HY     |  High Street/Castle Street
Perth                 |   S01011939       |  PH2 8PA     |  South Street
Falkirk and Stirling  |   S01013067       |  FK8 2LJ     |  Port Street

Kavanagh, Lee & Pryce (2016) produced a sensitivity analysis to assess the dependence of their results on the choice of centre, indicating they were robust to this choice. Alternative datazone centres could be used to explore the effect of definition for these analyses too.

### Linking 2001 to 2011 datazones

Income deprived and total population counts were reweighted from 2001 to 2011 datazone boundaries using a lookup file developed by [Paul Norman](http://www.geog.leeds.ac.uk/people/p.norman). This allowed weights to be produced, based on census output area overlap, allowing weighted 2011 population count estimates to be produced from counts enumerated at 2001 datazone level, based on the the level of overlap between 2001 and 2011 datazones. Further details about the lookup file was produced are available from [Paul Norman](p.d.norman@leeds.ac.uk) and the code used to perform the reweighting is reproduced below.

```{r 2011_reweight}
lkup <-  read_csv("data/paul_norman_file/paul_norman_dz2011_table.csv")
simd_combined <- read_csv("data/simd/simd_combined.csv")

lkup

simd_combined

lkup  %>%
  select(dz_2001, dz_2011)  %>%
  arrange(dz_2001, dz_2011)  %>%
  group_by(dz_2001, dz_2011)  %>%
  tally  %>% # produce n, giving number of OAs which contain particular groupings of dz_2001 and dz_2011
  arrange(dz_2011, dz_2001)  %>%
  select(dz_2011, dz_2001, n)  %>%
  group_by(dz_2011)   %>%
  arrange(dz_2011, dz_2001) %>%
  mutate(weight = n / sum(n)) -> weighting
weighting

weighting %>%  # Weighting by dz_2011
  left_join(simd_combined, by = c("dz_2001" = "datazone"))  %>%
  select(-simd_rank)  %>%  # Not meaningful to reweight rank
  group_by(dz_2011, year) %>%
  summarise_each( ~ sum(. * weight), 6:9) -> simd_2011_reweighted

simd_2011_reweighted

```

# Results

## Maps

Below is a map of Scotland in which each datazone has been coloured according to the nearest centre.

```{r show_map, eval = TRUE, echo = FALSE, cache = TRUE}
dz_2011 %>%
 append_data(., centre_distance, key.shp = "DataZone", key.data = "datazone", ignore.duplicates = T) %>%
 tm_shape(.) +
 tm_polygons(col = "nearest_centre", border.alpha = 0) -> map_nearest_centre

map_nearest_centre

```

This map shows the proportion of the population who are classified as income deprived for each SIMD
```{r show_id_map, echo = FALSE}
simd_2011_reweighted %>% 
  select(dz_2011, year, pop_incomedeprived, pop_total) %>%
  mutate(prop_id = pop_incomedeprived / pop_total) %>% 
  select(dz_2011, year, prop_id) %>% 
  spread(year, prop_id) %>% 
  append_data(
    shp = dz_2011, data = . ,
    key.shp = "DataZone", key.data = "dz_2011", ignore.na = T
  ) %>% 
  tm_shape(.) + 
  tm_facets(ncol = 4) + 
  tm_fill(col = c("2004", "2006", "2009", "2012"))

```

# Discussion

## Substantive Findings

## Limitations and recommendations for further research

Alternative methods of reweighting SIMD values released at 2001 datazone levels could be explored, including those which make use of more sophisticated GIS-based methods to apportion values to 2011 datazones, and of approaches which use SIMD component values available at a higher level of spatial disaggregation (e.g. output areas) to recalculate and reapportion SIMD scores to 2011 datazone geographies. As with all analyses, the additional benefit of alternative and more sophisticated approaches should be weighted against the additional time cost involved, and a judgement made about whether the ratio of marginal benefits to marginal costs represents a cost effective use of researcher time.

The robustness of the results to the choice of city centre could be explored in a sensitivity analysis.

### 2016-08-02 Script.R

@@ -5,484 +5,22 @@ require(pacman)
 pacman::p_load(
   readr, readxl,
   purrr, stringr, tidyr, dplyr, 
+  broom,
   rgeos,
   ggplot2, cowplot,
   tmap
 )
 
-
-# Load SIMD scores --------------------------------------------------------
-
-
-simd_combined <- read_csv("data/simd/simd_combined.csv")
-
-
-#Tasks 
-# Visualise choropleth for 2001
-# Visualise choropleth for 2011 
-# Visualise cartogram for 2001
-# Visualise cartogram for 2011 
-
-# Visualise both boundary layers for 2001 and 2011 
-
-
-# Load shapefiles  --------------------------------------------------------
-
-
-# 2001 datazone shapefiles ------------------------------------------------
-
-
-dz_2001 <- read_shape(file = "shapefiles/scotland_2001_datazones/scotland_dz_2001.shp")
-# To visualise this
-dz_2001 %>% 
-  tm_shape(.) + 
-  tm_borders() -> dz_2001_map  
-# To see as interactive Leaflet display
-#tmap_leaflet(dz_2001_map)
-  
-# And for 2011 datazones
-dz_2011 <- read_shape(file = "shapefiles/SG_DataZoneBdry_2011/SG_DataZone_Bdry_2011.shp")
-dz_2011 %>% 
-  tm_shape(.) + 
-  tm_borders( lty="dashed") -> dz_2011_map
-# Interactive
-#tmap_leaflet(dz_2011_map)
-
-
-# Both combined
-#tmap_leaflet(dz_2001_map + dz_2011_map)
-
-
-# To show simd_scores 
-
-
-simd_combined %>% 
-  select(datazone, year, simd_score) %>%
-  spread(year, simd_score) %>% 
-  append_data(
-    shp = dz_2001, data = . ,
-    key.shp = "zonecode", key.data = "datazone"
-              ) %>% 
-  tm_shape(.) + 
-  tm_facets(ncol = 4) + 
-  tm_fill(col = c("2004", "2006", "2009", "2012")) -> simd_chor_dz2001
-
-
-#   # Which of course says don't live in Glasgow...
-# save_tmap(simd_chor_dz2001, filename = "maps/dz2001_simd_chor.png", width = 40, height = 15,
-#           units = "cm", dpi = 300)
-
-
-# # Create cartograms 
-
-dz_2001_cart <- read_shape(
[…]

### 2016-08-02 scripts/do_dz2011_data_management.R

@@ -1,0 +1,72 @@
+
+# SIMD combined 
+
+simd_combined <- read_csv("data/simd/simd_combined.csv")
+# Paul Norman Lookup
+lkup <-  read_csv("data/paul_norman_file/paul_norman_dz2011_table.csv")
+
+
+# Produce reweighted SIMD scores
+lkup  %>% 
+  select(dz_2001, dz_2011)  %>% 
+  arrange(dz_2001, dz_2011)  %>%  
+  group_by(dz_2001, dz_2011)  %>% 
+  tally  %>% # produce n, giving number of OAs which contain particular groupings of dz_2001 and dz_2011
+  arrange(dz_2011, dz_2001)  %>% 
+  select(dz_2011, dz_2001, n)  %>% 
+  group_by(dz_2011)   %>% 
+  arrange(dz_2011, dz_2001) %>% 
+  mutate(weight = n / sum(n))   %>% # Weighting by dz_2011
+  left_join(simd_combined, by = c("dz_2001" = "datazone"))  %>% 
+  select(-simd_rank)  %>%  # Not meaningful to reweight rank
+  group_by(dz_2011, year) %>% 
+  summarise_each( ~ sum(. * weight, na.rm =T), 6:9) -> simd_2011_reweighted
+
+
+# Shapefile for 2011
+dz_2011 <- read_shape(file = "shapefiles/SG_DataZoneBdry_2011/SG_DataZone_Bdry_2011.shp")
+
+#qtm(dz_2011_shp)
+
+# 2011 centroids
+ttwa_centroids <- c(
+  Aberdeen = "S01006646",
+  Glasgow = "S01010265",
+  Edinburgh = "S01008677",
+  Dundee = "S01007705",
+  Inverness = "S01010620",
+  Perth = "S01011939",
+  `Falkirk and Stirling` = "S01013067"
+)
+
+# using rgeos::gCentroid
+
+calc_distance_to_centres <- function(shp, code_centre){
+  gCentroid(shp, byid = T) %>% 
+    as(., "data.frame") -> tmp
+  
+  data_frame(dz = as.character(shp@data$DataZone), x = tmp$x, y = tmp$y) %>% 
+    mutate(centre = dz == code_centre) %>% 
+    mutate(distance_to_centre = ((x - x[centre])^2 + (y - y[centre])^2)^0.5) -> output
+  output
+}
+
+
+fn <- function(val, nm){
+  dz_2011 %>% 
+    calc_distance_to_centres(., val) %>% 
[…]

### 2016-08-02 scripts/produce_dz2011_regression_bubble.R

@@ -1,0 +1,116 @@
+# Regressions -------------------------------------------------------------
+
+
+# The purpose of this section is 
+# 1) produce a comparable estimate of the gradient of 
+# change in poverty against distance for each place 
+# 2) produce gradients of change in total population 
+# 3) produce gradients of change in income deprived population 
+
+manage_propid_regression <- function(this_city){
+  simd_2011_reweighted  %>% 
+    ungroup() %>%  
+    left_join(centre_distance, by = c("dz_2011" = "datazone")) %>% 
+    filter(nearest_centre == this_city) %>%
+    distinct() %>% 
+    select(dz_2011, distance_to_centre, year, pop_incomedeprived, pop_total) %>% 
+    mutate(distance_to_centre = distance_to_centre / 1000) %>% 
+    mutate(prop_id = pop_incomedeprived / pop_total) %>% 
+    filter(distance_to_centre >= 1.0, distance_to_centre < 12.0) %>% 
+    select(dz_2011, year, distance_to_centre, prop_id) %>% 
+    spread(year, prop_id) %>% 
+    mutate(change = `2012` - `2004`) %>% 
+    lm(change ~ distance_to_centre, .) %>% 
+    tidy() %>% 
+    mutate(place = this_city) %>% 
+    select(place, everything())
+}
+
+c("Aberdeen", "Dundee", "Edinburgh", "Glasgow") %>% 
+  map_df(.f = manage_propid_regression) -> propid_regression
+
+manage_numerator_regression <- function(this_city){
+  simd_2011_reweighted  %>% 
+    ungroup() %>%  
+    left_join(centre_distance, by = c("dz_2011" = "datazone")) %>% 
+    filter(nearest_centre == this_city) %>%
+    distinct() %>% 
+    select(dz_2011, distance_to_centre, year, pop_incomedeprived) %>% 
+    mutate(distance_to_centre = distance_to_centre / 1000) %>% 
+    filter(distance_to_centre >= 1.0, distance_to_centre < 12.0) %>% 
+    select(dz_2011, year, distance_to_centre, pop_incomedeprived) %>% 
+    spread(year, pop_incomedeprived) %>% 
+    mutate(`2012` = `2012` + 0.5, `2004` = `2004` + 0.5) %>% # Continuity correction
[…]

### 2016-08-02 scripts/results_for_2001dz.R

@@ -1,0 +1,411 @@
+rm(list = ls())
+
+require(pacman)
+
+pacman::p_load(
+  readr, readxl,
+  purrr, stringr, tidyr, dplyr, 
+  rgeos,
+  ggplot2, cowplot,
+  tmap
+)
+
+
+# Load SIMD scores --------------------------------------------------------
+
+
+simd_combined <- read_csv("data/simd/simd_combined.csv")
+
+
+#Tasks 
+# Visualise choropleth for 2001
+# Visualise choropleth for 2011 
+# Visualise cartogram for 2001
+# Visualise cartogram for 2011 
+
+# Visualise both boundary layers for 2001 and 2011 
+
+
+# Load shapefiles  --------------------------------------------------------
+
+
+# 2001 datazone shapefiles ------------------------------------------------
+
+
+dz_2001 <- read_shape(file = "shapefiles/scotland_2001_datazones/scotland_dz_2001.shp")
+# To visualise this
+dz_2001 %>% 
+  tm_shape(.) + 
+  tm_borders() -> dz_2001_map  
+# To see as interactive Leaflet display
+#tmap_leaflet(dz_2001_map)
+  
+# And for 2011 datazones
+dz_2011 <- read_shape(file = "shapefiles/SG_DataZoneBdry_2011/SG_DataZone_Bdry_2011.shp")
+dz_2011 %>% 
+  tm_shape(.) + 
+  tm_borders( lty="dashed") -> dz_2011_map
+# Interactive
+#tmap_leaflet(dz_2011_map)
+
+
+# Both combined
+#tmap_leaflet(dz_2001_map + dz_2011_map)
+
+
+# To show simd_scores 
+
+
+simd_combined %>% 
+  select(datazone, year, simd_score) %>%
+  spread(year, simd_score) %>% 
+  append_data(
+    shp = dz_2001, data = . ,
+    key.shp = "zonecode", key.data = "datazone"
+              ) %>% 
+  tm_shape(.) + 
+  tm_facets(ncol = 4) + 
+  tm_fill(col = c("2004", "2006", "2009", "2012")) -> simd_chor_dz2001
+
+
+#   # Which of course says don't live in Glasgow...
+# save_tmap(simd_chor_dz2001, filename = "maps/dz2001_simd_chor.png", width = 40, height = 15,
+#           units = "cm", dpi = 300)
+
+
+# # Create cartograms 
+
[…]

### 2016-08-02 scripts/do_dz2011_data_management.R
save data management script

@@ -26,6 +26,8 @@ lkup  %>%
 # Shapefile for 2011
 dz_2011 <- read_shape(file = "shapefiles/SG_DataZoneBdry_2011/SG_DataZone_Bdry_2011.shp")
 
+# Cartogram for 2011 
+dz_2011_cart <- read_shape(file = "shapefiles/dz_2011_cart/dz_2011_cart.shp", current.projection = "longlat")
 #qtm(dz_2011_shp)
 
 # 2011 centroids

### 2016-08-03 scripts/from_gavin/RCI.R
add script from Gavin

@@ -1,0 +1,55 @@
+RCI <- function(povvec, popvec, order)
+{
+    
+    #Storage vectors for X = ordered vector of people "in poverty"; Y = "not in pov"
+    X <- 1:length(povvec)
+    Y <- X
+    
+    #povvec[i] is the number of people in poverty (on JSA, Income Support or whatever) in area i
+    #X[i] is total number of people in poverty in the i units closest to city centre,
+    #divided by overall number of people in poverty.
+    for(i in 1:length(X))
+    {
+        X[i] <- sum(povvec[order[1:i]])/sum(povvec)
+        Y[i] <- sum(popvec[order[1:i]] - povvec[order[1:i]])/sum(popvec - povvec)
+    }
+    
+    #Variable to hold result
+    res.rce = 0
+    for(i in 2:length(X))
+    {
+        res.rce <- res.rce + X[i-1] * Y[i] - X[i] * Y[i-1]
+    }
+    
+    return(res.rce)
+}
+
+
+
+
+
+RCI.path <- function(povvec, popvec, order, K)
+{
+    
+    #Storage vectors for X = ordered vector of people "in poverty"; Y = "not in pov"
+    X <- 1:length(povvec)
+    Y <- X
+    
+    #povvec[i] is the number of people in poverty (on JSA, Income Support or whatever) in area i
+    #X[i] is total number of people in poverty in the i units closest to city centre,
+    #divided by overall number of people in poverty.
+    for(i in 1:K)
+    {
+        X[i] <- sum(povvec[order[1:i]])/sum(povvec[order[1:K]])
+        Y[i] <- sum(popvec[order[1:i]] - povvec[order[1:i]])/sum(popvec[order[1:K]] - povvec[order[1:K]])
+    }
+    
+    #Variable to hold result
+    res.rce = 0
+    for(i in 2:K)
+    {
+        res.rce <- res.rce + X[i-1] * Y[i] - X[i] * Y[i-1]
+    }
+    
+    return(res.rce)
+}

### 2016-08-03 markdown/draft.RmD
save markdown doc with its gubbins

@@
 
-```{r setup01, include=TRUE, echo = TRUE}
+```{r setup01, include=TRUE, echo = FALSE, message = FALSE, warning = FALSE}
 rm(list = ls())
@@
 library(knitr)
-getwd()
+
 opts_knit$set(root.dir=normalizePath('../'))
@@
 
-```{r setup02, include = TRUE, cache = FALSE, echo = TRUE}
-getwd()
+```{r setup02, include = TRUE, cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE}
+
 require(pacman)
@@
   purrr, stringr, tidyr, dplyr,
+  broom,
   rgeos,
   ggplot2, cowplot,
-  tmap
+  tmap, RColorBrewer
 )
@@
 source("scripts/do_dz2011_data_management.R")
-ls()
-centre_distance
 
@@
 
-And now here is a third chunk to check 1) that the working directory has not been reset 
- and 2) to see if the centre_distance object still exists
-
 # Introduction
 
-Here is the contents of centre_distance
+This is the introduction section to the report.
 
@@
 The briefing paper defined 'poverty' in three ways, each relating to specific UK state benefit receipts for persons of working age:
+
 1. The proportion of local populations receiving Income Support (IS)
@@
 The purposes of this briefing paper are two-fold:
+
 1.
@@
 
-The SIMD is ...
+The SIMD is used by the Scottish Government to identify places suffering from deprivation.
+Different editions of the SIMD have been produced in 2004, 2006, 2012, with a new release scheduled for August 2016, and made available at the small area known as the datazone, equivalent to the Lower Super Output Area (LSOA) in England & Wales.
+The methodology differs slightly from one edition to the next, but for all releases estimates of the number of individuals within each datazone who are 'income deprived', as well as overall and domain-specific scores and rankings for deprivation.
+The definition of 'income deprived' varies between household types and age groups, to account for the different needs of different demographic groups and the efficiency gains associated with living in multi-person rather than single person households.
 SIMD data releases for 2004, 2006, 2012 and 2016 were downloaded from the [SIMD website](http://www.gov.scot/Topics/Statistics/SIMD/DataAnalysis).
 The total population counts, and the population count of the subpopulation defined as 'income deprived', were extracted from each release and combined in a single dataset.
+The 'income deprived' and 'total' population counts were the focus of this report as the RCI measure, like many other segregation measures, involves comparing two mutually exclusive groups.
 
@@
 
+The majority of the results presented will be for the four cities Aberdeen, Dundee, Edinburgh and Glasgow, as these are the largest and most centralised of the areas considered.
+Within this Glasgow is historically the most populous, has the highest population density, and has the highest concentration of deprived areas.
+
 ### Linking 2001 to 2011 datazones
@@
 Further details about the lookup file was produced are available from [Paul Norman](p.d.norman@leeds.ac.uk) and the code used to perform the reweighting is reproduced below.
+
+# Results
+
+## Maps
+
+Below is a map of Scotland in which each datazone has been coloured according to the nearest centre.
+
+```{r show_map, eval = TRUE, echo = FALSE, cache = FALSE, message = FALSE, warning = FALSE}
+dz_2011 %>%
+ append_data(., centre_distance, key.shp = "DataZone", key.data = "datazone", ignore.duplicates = T) %>%
+ tm_shape(.) +
+ tm_polygons(col = "nearest_centre", border.alpha = 0) -> map_nearest_centre
+
+map_nearest_centre
+
+```
+
+As this map indicates places in the Orkney and Shetland Islands, and elsewhere in the remote and rural highlands and islands, are arbitrily assigned to either Aberdeen or Inverness, even though they are not part of any meaningful functional urban geography.
+However these are easily identifiable through their absolute distance from either centre, and are not part of the main analysis.
+
+The following map shows the proportion of the population who are classified as income deprived for each datazone, in the years 2004, 2006, 2009 and 2012.
+Given that so much of Scotland is rural, few areas appear income deprived using a standard choropleth.
+However, patterns of high income deprivation are apparent in and near Glasgow.
+
+```{r show_id_map, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE}
+simd_2011_reweighted %>% 
+  select(dz_2011, year, pop_incomedeprived, pop_total) %>%
+  mutate(prop_id = pop_incomedeprived / pop_total) %>% 
+  mutate(prop_id = ifelse(prop_id > 0.5, 0.5, prop_id)) %>% 
+  select(dz_2011, year, prop_id) %>% 
+  spread(year, prop_id) %>% 
+  append_data(
+    shp = dz_2011, data = . ,
+    key.shp = "DataZone", key.data = "dz_2011", ignore.na = T
+  ) %>% 
+  tm_shape(.) + 
+  tm_facets(ncol = 2) + 
+  tm_fill(
+    col = c("2004", "2006", "2009", "2012"), 
+    palette = brewer.pal(10, "Spectral"), 
+    breaks = seq(0, 0.5, by = 0.05), showNA = F
+    ) -> id_chor_dz2011
+
+id_chor_dz2011
+```
+
+The map below focuses in on income deprivation within Glasgow, and in particular within datazones not more than 15 km from Glasgow city centre.
+
+```{r show_id_map_glasgow, echo = F, eval = T, message = FALSE, warning = FALSE}
+
+simd_2011_reweighted %>% 
+  select(dz_2011, year, pop_incomedeprived, pop_total) %>%
+  mutate(prop_id = pop_incomedeprived / pop_total) %>% 
+  mutate(prop_id = ifelse(prop_id > 0.5, 0.5, prop_id)) %>% 
+  select(dz_2011, year, prop_id) %>% 
+  left_join(centre_distance, by = c("dz_2011" = "datazone")) %>% 
+  filter(nearest_centre == "Glasgow") %>% 
+  filter(distance_to_centre < 15000) %>% 
+  select(dz_2011, year, prop_id) %>% 
+  spread(year, prop_id) %>% 
+  append_data(
+    shp = dz_2011, data = . ,
+    key.shp = "DataZone", key.data = "dz_2011", ignore.na = T
+  ) -> shp_jn
+
+shp_jn <- shp_jn[!is.na(shp_jn$`2004`), ]
+shp_jn %>% 
+  tm_shape(.) + 
+  tm_facets(ncol = 2) + 
+  tm_fill(
+    col = c("2004", "2006", "2009", "2012"), 
+    palette = brewer.pal(10, "Spectral"), 
+    breaks = seq(0, 0.5, by = 0.05), showNA = F
+    ) +
+  tm_legend(legend.show = F) -> id_chor_dz2011_glasgow
+
+id_chor_dz2011_glasgow
+
+```
+
+From this map there are suggestions of a decentralisation of poverty, at least as defined as income deprived, in particular between 2009 and 2012.
+However there are also clear pockets of areas, in the south, east and north west of Glasgow city, with persistently high levels of income deprivation.
+
+An equivalent map for Edinburgh is produced below:
+
+```{r show_id_map_edinburgh, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE}
+
+simd_2011_reweighted %>% 
+  select(dz_2011, year, pop_incomedeprived, pop_total) %>%
+  mutate(prop_id = pop_incomedeprived / pop_total) %>% 
+  mutate(prop_id = ifelse(prop_id > 0.5, 0.5, prop_id)) %>% 
+  select(dz_2011, year, prop_id) %>% 
+  left_join(centre_distance, by = c("dz_2011" = "datazone")) %>% 
+  filter(nearest_centre == "Edinburgh") %>% 
+  filter(distance_to_centre < 15000) %>% 
+  select(dz_2011, year, prop_id) %>% 
+  spread(year, prop_id) %>% 
+  append_data(
+    shp = dz_2011, data = . ,
+    key.shp = "DataZone", key.data = "dz_2011", ignore.na = T
+  ) -> shp_jn
+
+shp_jn <- shp_jn[!is.na(shp_jn$`2004`), ]
+shp_jn %>% 
+  tm_shape(.) + 
+  tm_facets(ncol = 2) + 
+  tm_fill(
+    col = c("2004", "2006", "2009", "2012"), 
+    palette = brewer.pal(10, "Spectral"), 
+    breaks = seq(0, 0.5, by = 0.05), showNA = F
+    ) +
+  tm_legend(legend.show = F) -> id_chor_dz2011_edinburgh
+
+id_chor_dz2011_edinburgh
+
+```
+
+This map also suggests that, as in Glasgow, drivers of increasing levels of decentralisation of poverty in Edinburgh include both rapid falls in poverty in the centre, alongside some more suburban areas with persistently high levels of poverty.
+In Edinburgh's case, in pockets towards the West, North West, and South/South East of the city centre.
+
+The equivalent map for Dundee is shown below:
+
+```{r show_id_map_dundee, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE}
+
+simd_2011_reweighted %>% 
+  select(dz_2011, year, pop_incomedeprived, pop_total) %>%
+  mutate(prop_id = pop_incomedeprived / pop_total) %>% 
+  mutate(prop_id = ifelse(prop_id > 0.5, 0.5, prop_id)) %>% 
+  select(dz_2011, year, prop_id) %>% 
+  left_join(centre_distance, by = c("dz_2011" = "datazone")) %>% 
+  filter(nearest_centre == "Dundee") %>% 
+  filter(distance_to_centre < 15000) %>% 
+  select(dz_2011, year, prop_id) %>% 
+  spread(year, prop_id) %>% 
+  append_data(
+    shp = dz_2011, data = . ,
+    key.shp = "DataZone", key.data = "dz_2011", ignore.na = T
+  ) -> shp_jn
+
+shp_jn <- shp_jn[!is.na(shp_jn$`2004`), ]
+shp_jn %>% 
+  tm_shape(.) + 
+  tm_facets(ncol = 2) + 
+  tm_fill(
+    col = c("2004", "2006", "2009", "2012"), 
+    palette = brewer.pal(10, "Spectral"), 
+    breaks = seq(0, 0.5, by = 0.05), showNA = F
+    ) +
+  tm_legend(legend.show = F) -> id_chor_dz2011_dundee
+
+id_chor_dz2011_dundee
+
+```
+
+Within Dundee the patterns of spatial persistence of poverty appear most consistent over time, with a clearer demarcation between urban and non-urban areas.
+The functional urban geography of Dundee appears more spatially concentrated than for Glasgow and Edinburgh.
+
+The equivalent map for Aberdeen, historically the most affluent and rural of the four cities, is shown here:
+
+```{r show_id_map_aberdeen, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE}
+
+simd_2011_reweighted %>% 
+  select(dz_2011, year, pop_incomedeprived, pop_total) %>%
+  mutate(prop_id = pop_incomedeprived / pop_total) %>% 
+  mutate(prop_id = ifelse(prop_id > 0.5, 0.5, prop_id)) %>% 
+  select(dz_2011, year, prop_id) %>% 
+  left_join(centre_distance, by = c("dz_2011" = "datazone")) %>% 
+  filter(nearest_centre == "Aberdeen") %>% 
+  filter(distance_to_centre < 15000) %>% 
+  select(dz_2011, year, prop_id) %>% 
+  spread(year, prop_id) %>% 
+  append_data(
+    shp = dz_2011, data = . ,
+    key.shp = "DataZone", key.data = "dz_2011", ignore.na = T
+  ) -> shp_jn
+
+shp_jn <- shp_jn[!is.na(shp_jn$`2004`), ]
+shp_jn %>% 
+  tm_shape(.) + 
+  tm_facets(ncol = 2) + 
+  tm_fill(
+    col = c("2004", "2006", "2009", "2012"), 
+    palette = brewer.pal(10, "Spectral"), 
+    breaks = seq(0, 0.5, by = 0.05), showNA = F
+    ) +
+  tm_legend(legend.show = F) -> id_chor_dz2011_aberdeen
+
+id_chor_dz2011_aberdeen
+
+```
+
+As with Glasgow and Edinburgh, it appears both that the greatest changes have occurred between 2009 and 2012, and also that any fall in RCI scores are likely due to persistence of poverty in some local areas, with a high income deprivation area clearly visible in all years to the north west of the city centre.
+
+## Plots of the distance/deprivation relationship
+
+In the figure below, the vertical axis shows the proportion change in income deprivation from 2004 to 2012, and the horizontal axis shows the distance of datazones from the city centres, plotted on a log scale, for each of the four main Scottish cities.
+Individual datazones are indicated with faint circles, and a nonlinear smoother has been added showing how this the change in income deprivation varies as a function of distance.
+A dashed horizontal line has been added indicating no change in deprivation between the two period, and vertical lines added at 1km and 12km, as below 1km there are few observations, and above 12km the effect of a functional urban geography appears small.
+
+```{r show_change_in_id, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+decent_plot_for_city <- function(this_city){
+  simd_2011_reweighted  %>% 
+    ungroup() %>%  
+    left_join(centre_distance, by = c("dz_2011" = "datazone")) %>% 
+    filter(nearest_centre == this_city) %>%
+    distinct() %>% 
+    mutate(prop_id = pop_incomedeprived / pop_total) %>% 
+    select(dz_2011, distance_to_centre, year, prop_id) %>% 
+    mutate(distance_to_centre = distance_to_centre / 1000) %>% 
+    spread(year, prop_id) %>% 
+    mutate(change = `2012` - `2004`) %>% 
+    ggplot(., aes(x = distance_to_centre, y = change)) + 
+    geom_point( alpha = 0.01) + 
+    scale_x_log10(limits = c(0.5, 50), breaks = c(0.5, 1, 2, 5, 10, 20, 50)) + 
+    scale_y_continuous(limits = c(-0.25, 0.25) ) +
+    stat_smooth() + 
+    geom_hline(aes(yintercept = 0), linetype = "dashed") + 
+    geom_vline(aes(xintercept = 1e+0)) + 
+    geom_vline(aes(xintercept = 1.2e+1)) + 
+    labs(x = "Distance to centre (km)", y = "Change in ID",
+         title = this_city)
+}
+
+plot_grid(
+  decent_plot_for_city("Aberdeen"),
+  decent_plot_for_city("Dundee"),
+  decent_plot_for_city("Edinburgh"),
+  decent_plot_for_city("Glasgow"),
+  nrow = 2
+)
+
+```
+
+These plots suggest a qualitatively similar relationship between change in deprivation and distance, up to around 12km from city centre, for Aberdeen, Edinburgh and Glasgow, with Glasgow showing the strongest gradient of change between these two periods.
+Dundee shows a simpilar kind of relationship, but concentrated within the first 5km, rather than 12km, of distance from city centre.
+This is consistent with the maps of income deprivation in Dundee, showing much greater levels of spatial concentration than for the other cities.
+
+### Numerator and denominator changes
+
+The falls in income deprivation nearer city centres are a function both of changing numerators - i.e.
+changes to income deprived population sizes - and changing denominators - i.e.
+changes to total population sizes.
+To gain a better understanding of the relative influence of these two changes to the relationship shown in the previous figure, the next two figures show the proportionate change from 2004 to 2012 in income deprived populations, and total populations.
+
+The next figure shows how the income deprived population has changed between the two periods.
+
+```{r plot_changeidpop, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+idpopchange_plot_for_city <- function(this_city){
+  simd_2011_reweighted  %>% 
+    ungroup() %>%  
+    left_join(centre_distance, by = c("dz_2011" = "datazone")) %>% 
+    filter(nearest_centre == this_city) %>%
+    distinct() %>% 
+    select(dz_2011, distance_to_centre, year, pop_incomedeprived) %>% 
+    mutate(distance_to_centre = distance_to_centre / 1000) %>% 
+    spread(year, pop_incomedeprived) %>% 
+    mutate(change = (`2012` / `2004`) - 1) %>% 
+    ggplot(., aes(x = distance_to_centre, y = change)) + 
+    geom_point( alpha = 0.01) + 
+    scale_x_log10(limits = c(0.5, 50), breaks = c(0.5, 1, 2, 5, 10, 20, 50)) + 
+    scale_y_continuous(limits = c(-.5, .5)) +    
+    stat_smooth() + 
+    geom_vline(aes(xintercept = 1e+0)) + 
+    geom_vline(aes(xintercept = 1.2e+1)) +
+    geom_hline(aes(yintercept = 0), linetype = "dashed") + 
+    labs(x = "Distance to centre (km)", y = "Change in ID pop",
+         title = this_city)
+}
+
+plot_grid(
+  idpopchange_plot_for_city("Aberdeen"),
+  idpopchange_plot_for_city("Dundee"),
+  idpopchange_plot_for_city("Edinburgh"),
+  idpopchange_plot_for_city("Glasgow"),
+  nrow = 2
+)
+```
+
+This plot indicates proportionate falls in the ID population shares at distances closes to the city centre, up to around 20% in Aberdeen and Glasgow, and either proportionate rises (Dundee and Edinburgh) or little change in proportions (Aberdeen and Glasgow) at greater distance from the city centre.
+This seems to suggest, for Aberdeen and Glasgow in particular, that 'gentrification of the centre' could be as much or more of an explanation for falls in RCI as 'impoverishment of the periphery', though that there are somewhat different socioeconomic mechanisms in place for the four cities.
+
+The following plot shows the change in the total population size as a function of distance from city centre.
+(The smoother line at distances less than 1km should likely be ignored for most cities as it involves extrapolation based on very few observations).
+```{r plot_changetotalpop, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+totalpopchange_plot_for_city <- function(this_city){
+  simd_2011_reweighted  %>% 
+    ungroup() %>%  
+    left_join(centre_distance, by = c("dz_2011" = "datazone")) %>% 
+    filter(nearest_centre == this_city) %>%
+    distinct() %>% 
+    select(dz_2011, distance_to_centre, year, pop_total) %>% 
+    mutate(distance_to_centre = distance_to_centre / 1000) %>% 
+    spread(year, pop_total) %>% 
+    mutate(change = (`2012` / `2004`) - 1) %>% 
+    ggplot(., aes(x = distance_to_centre, y = change)) + 
+    geom_point( alpha = 0.01) + 
+    scale_x_log10(limits = c(0.5, 50), breaks = c(0.5, 1, 2, 5, 10, 20, 50)) + 
+    scale_y_continuous(limits = c(-.5, .5)) +    
+    stat_smooth() + 
+    geom_vline(aes(xintercept = 1e+0)) + 
+    geom_vline(aes(xintercept = 1.2e+1)) +
+    geom_hline(aes(yintercept = 0), linetype = "dashed") + 
+    labs(x = "Distance to centre (km)", y = "Change in total pop",
+         title = this_city)
+}
+
+plot_grid(
+  totalpopchange_plot_for_city("Aberdeen"),
+  totalpopchange_plot_for_city("Dundee"),
+  totalpopchange_plot_for_city("Edinburgh"),
+  totalpopchange_plot_for_city("Glasgow"),
+  nrow = 2
+)
+```
+This plot suggests that there have been substantial increases in population sizes in Aberdeen, Edimburgh and Glasgow over this period, in datazones closer to the city centre, with the greatest changes occurring near the centre of Glasgow.
+In Aberdeen, and to a lesser extent for Dundee, there have also been increases in population shares living at distances of 12km and more from the city, suggesting greatest consistency in population sizes at intermediate distances.
+
+The following plot attempts to summarise the previous series of results, in order the represent how both total population sizes and income deprived population sizes have changed with distance, and the overall strength of the relationship between income deprived population share and distance from city centre.
+
+For each of the four cities, separate linear regressions were produced to extract the gradient of changes in total population size, income deprived population size, and changes in income deprived population shares with distance, over distances between 1km and 12km.
+The horizontal axis shows the gradient of change in total population with distance, the vertical axis shows graident of change in income deprived population with distance.
+Each city is represented by a separate point, labelled G for Glasgow, A for Aberdeen, D for Dundee and E for Edinburgh.
+The darkness of these symbols indicates the magnitude of the overall change in the income deprived population shares as a function of distance, with darker shades indicating a larger relationship and lighter shades a smaller relationship.
+
+```{r plot_all_bubble, eval = TRUE, echo = FALSE , message = FALSE, warning = FALSE}
+
+source("scripts/produce_dz2011_regression_bubble.R")
+
+# As a scatterplot?
+
+simple_results %>% 
+  mutate(place2 = str_sub(place, 1, 1)) %>% 
+  select(type, place2, estimate) %>% 
+  spread(type, estimate) %>% 
+  ggplot(., aes(x = denominator, y = numerator)) + 
+  geom_point(size = 10, colour = "lightblue") +
+  geom_text(aes(label = place2, alpha =  propid), show.legend = F) +
+  coord_cartesian(xlim = c(-0.1, 0.1), ylim = c(0.0, 0.1)) +
+  geom_hline(yintercept = 0, linetype = "dashed") + 
+  geom_vline(xintercept = 0, linetype = "dashed") + 
+  labs(x = "change in total population with distance", y = "change in income deprived population with distance")
+
+```
+
+## RCI calculations
+
+This section will present the results of the RCI calculations, i.e. more formal attempt to replicate the results of Kavanagh, Pryce & Lee (2016) using a different indicator and data source.
+
+# Discussion
+
+## Substantive Findings
+
+The apparent fall in income deprivation proportions in more central areas of cities, in particular Glasgow, and in particular the fact the fall appeared greatest between 2008 and 2012, suggests there may be issues with the way in which income deprivation has been defined, and how effectively or otherwise this connects to broader and alternative definitions of poverty other this period.
+Although the definition of income deprived is not completely consistent from one SIMD release to the next, they tend to be based on claims for particular types of state benefit.
+During the 2008 to 2012 period the level of conditionality associated with many benefits increased for people of working age, while at the same time the longest sustained fall and plateau in real wages in over half a century occurred.
+Unemployment rates did not rise as they had in previous recessions, but levels of self-employment, short-term contracts, and part time work all increased.
+The self-employed, in particular, are not eligible for many of the benefits that people earning low wages as employees receive, and so the fall the income deprivation may reflect this compositional change in the workforce.
+Similarly increasing casualisation of employment, with variable hours and pay from one week to the next, may put off many people earning low wages on average from applying for 'top up' benefits such as tax credits.
+Again, indications that there have been falls in Scotland in income deprivation between 2008 to 2012 suggest statistical artefacts.
+
+## Limitations and recommendations for further research
+
+Alternative methods of reweighting SIMD values released at 2001 datazone levels could be explored, including those which make use of more sophisticated GIS-based methods to apportion values to 2011 datazones, and of approaches which use SIMD component values available at a higher level of spatial disaggregation (e.g.
+output areas) to recalculate and reapportion SIMD scores to 2011 datazone geographies.
+As with all analyses, the additional benefit of alternative and more sophisticated approaches should be weighted against the additional time cost involved, and a judgement made about whether the ratio of marginal benefits to marginal costs represents a cost effective use of researcher time.
+
+The robustness of the results to the choice of city centre could be explored in a sensitivity analysis.
+
+# Appendices
+
+## Appendix A: Code for reweighting SIMD available for 2001 datazones onto 2011 datazones
+
+Within the script below, the object lkup is the lookup file produced by Paul Norman, and simd_combined contains data from the 2004, 2006, 2009, 2012 and 2016 releases of the SIMD.
 
@@
 ```
-
-# Results
-
-## Maps
-
-Below is a map of Scotland in which each datazone has been coloured according to the nearest centre.
-
-```{r show_map, eval = TRUE, echo = FALSE, cache = TRUE}
-dz_2011 %>%
- append_data(., centre_distance, key.shp = "DataZone", key.data = "datazone", ignore.duplicates = T) %>%
- tm_shape(.) +
- tm_polygons(col = "nearest_centre", border.alpha = 0) -> map_nearest_centre
-
-map_nearest_centre
-
-```
-
-This map shows the proportion of the population who are classified as income deprived for each SIMD
-```{r show_id_map, echo = FALSE}
-simd_2011_reweighted %>% 
-  select(dz_2011, year, pop_incomedeprived, pop_total) %>%
-  mutate(prop_id = pop_incomedeprived / pop_total) %>% 
-  select(dz_2011, year, prop_id) %>% 
-  spread(year, prop_id) %>% 
-  append_data(
-    shp = dz_2011, data = . ,
-    key.shp = "DataZone", key.data = "dz_2011", ignore.na = T
-  ) %>% 
-  tm_shape(.) + 
-  tm_facets(ncol = 4) + 
-  tm_fill(col = c("2004", "2006", "2009", "2012"))
-
-```
-
-# Discussion
-
-## Substantive Findings
-
-## Limitations and recommendations for further research
-
-Alternative methods of reweighting SIMD values released at 2001 datazone levels could be explored, including those which make use of more sophisticated GIS-based methods to apportion values to 2011 datazones, and of approaches which use SIMD component values available at a higher level of spatial disaggregation (e.g.
-output areas) to recalculate and reapportion SIMD scores to 2011 datazone geographies.
-As with all analyses, the additional benefit of alternative and more sophisticated approaches should be weighted against the additional time cost involved, and a judgement made about whether the ratio of marginal benefits to marginal costs represents a cost effective use of researcher time.
-
-The robustness of the results to the choice of city centre could be explored in a sensitivity analysis.

### 2016-08-03 Script.R

@@ -81,24 +81,25 @@ decent_plot_for_city <- function(this_city){
     spread(year, prop_id) %>% 
     mutate(change = `2012` - `2004`) %>% 
     ggplot(., aes(x = distance_to_centre, y = change)) + 
-    geom_point( alpha = 0.1) + 
-    
-    scale_x_log10(breaks = c(0.5, 1, 2, 5, 10, 20, 50, 100)) + 
-    scale_y_continuous(limits = c(-1.0, 0.5) ) +
+    geom_point( alpha = 0.01) + 
+    scale_x_log10(limits = c(0.5, 50), breaks = c(0.5, 1, 2, 5, 10, 20, 50)) + 
+    scale_y_continuous(limits = c(-0.25, 0.25) ) +
     stat_smooth() + 
     geom_hline(aes(yintercept = 0), linetype = "dashed") + 
     geom_vline(aes(xintercept = 1e+0)) + 
     geom_vline(aes(xintercept = 1.2e+1)) + 
-    labs(x = "Distance to centre (km)", y = "Change in income deprived proportion",
+    labs(x = "Distance to centre (km)", y = "Change in ID",
          title = this_city)
 }
 
-decent_plot_for_city("Aberdeen")
-decent_plot_for_city("Dundee")
-decent_plot_for_city("Edinburgh")
-decent_plot_for_city("Glasgow")
 
-
+plot_grid(
+  decent_plot_for_city("Aberdeen"),
+  decent_plot_for_city("Dundee"),
+  decent_plot_for_city("Edinburgh"),
+  decent_plot_for_city("Glasgow"),
+  nrow = 2
+)
 # Change in population size 
 
 popchange_plot_for_city <- function(this_city){
@@ -160,10 +161,6 @@ idpopchange_plot_for_city("Glasgow")
 
 source("scripts/produce_dz2011_regression_bubble.R")
 
-simple_results %>% 
-  ggplot(., aes(y = place, group = type, x = estimate, shape = type)) + 
-  geom_point() +   
-  geom_vline(aes(xintercept = 0), linetype = "dashed")
 
 # As a scatterplot?
 

### 2016-08-04 markdown/draft.RmD
do RCI and update markdown doc many times

@@
 author: "Jon Minton"
-date: "2 August 2016"
+date: "4 August 2016"
 output: html_document
@@
 ```
+
+# Caveats
+
+* Work very much in progress
+* Please don't steal
 
@@
 Producing the results involved doing the following:
+
 * Defining centroids for each place.
@@
 
-### Linking 2001 to 2011 datazones
-
-Income deprived and total population counts were reweighted from 2001 to 2011 datazone boundaries using a lookup file developed by [Paul Norman](http://www.geog.leeds.ac.uk/people/p.norman).
-This allowed weights to be produced, based on census output area overlap, allowing weighted 2011 population count estimates to be produced from counts enumerated at 2001 datazone level, based on the the level of overlap between 2001 and 2011 datazones.
-Further details about the lookup file was produced are available from [Paul Norman](p.d.norman@leeds.ac.uk) and the code used to perform the reweighting is reproduced below.
-
-# Results
-
-## Maps
+### Apportioning nearest centre to datazones
 
@@
 However these are easily identifiable through their absolute distance from either centre, and are not part of the main analysis.
+
+### Linking 2001 to 2011 datazones
+
+Income deprived and total population counts were reweighted from 2001 to 2011 datazone boundaries using a lookup file developed by [Paul Norman](http://www.geog.leeds.ac.uk/people/p.norman).
+This allowed weights to be produced, based on census output area overlap, allowing weighted 2011 population count estimates to be produced from counts enumerated at 2001 datazone level, based on the the level of overlap between 2001 and 2011 datazones.
+Further details about the lookup file was produced are available from [Paul Norman](p.d.norman@leeds.ac.uk) and the code used to perform the reweighting is reproduced below.
+
+# Results
+
+## RCI calculations
+
+This section will present the results of the RCI calculations, i.e. more formal attempt to replicate the results of Kavanagh, Pryce & Lee (2016) using a different indicator and data source.
+
+For Glasgow, the RCI scores (with 95% credible intervals, CrIs) for 2004 and 2012, and for the difference between these two periods, are as follows:
+
+Quantile   | RCI, 2004 | RCI, 2012  | Change  
+---------- | --------  | ---------- | -------  
+median     | 0.144     | 0.093      | -0.051
+2.5% CrI   | 0.142     | 0.091      | -0.054
+97.5% CrI  | 0.146     | 0.095      | -0.048
+
+From this we can conclude that there has been a statistically significant fall in RCI over these two periods using the SIMD income deprivation measure, as there had been for out-of-work benefits-based measures for 2001 and 2011 in Kavanagh at al 2016.
+
+As with Kavanagh et al 2016 Dissimilarity index scores (D) were also calculated for both periods, along with change between them.
+These are shown in the table below, and also show statistically significant falls.
+
+Quantile   | D, 2004 | D, 2012  | Change  
+---------- | ------- | -------- | ------- 
+median     | 0.368   | 0.315    | -0.053
+2.5% CrI   | 0.366   | 0.313    | -0.056
+97.5% CrI  | 0.370   | 0.317    | -0.051
+
+## Maps
 
@@
 
+## Plotting temporal dependence
+
+The following plot shows scatterplots of the ID proportions in each Scottish datazone in 2004, 2009 and 2012.
+Within these plots the blue line indicates parity between values on the x and y axis, and the red dashed line shows a nonlinear smoother function between the two ID proportions being plotted, and hence the level of deviation between the actual relationship over the two years and parity.
+```{r show_temporal_dep, echo = FALSE, warning = FALSE, EVAL = TRUE}
+
+simd_2011_reweighted %>% 
+  filter(year %in% c(2004, 2012)) %>% 
+  mutate(prop_id = pop_incomedeprived / pop_total) %>% 
+  select(dz_2011, year, prop_id) %>% 
+  spread(year, prop_id) %>% 
+  rename(prop_id_2004 = `2004`, prop_id_2012 = `2012`) %>% 
+  ggplot(., aes(x = prop_id_2004, y = prop_id_2012)) + 
+  geom_point(alpha = 0.1) + 
+  geom_abline(colour = "blue", slope = 1, intercept = 0) + 
+  stat_smooth(colour = "red", linetype = "dashed") + 
+  labs(
+    x = "ID prop, 2004", y = "ID prop, 2012",
+    title = "2012 against 2004"
+       ) -> s_04_12
+
+simd_2011_reweighted %>% 
+  filter(year %in% c(2004, 2006)) %>% 
+  mutate(prop_id = pop_incomedeprived / pop_total) %>% 
+  select(dz_2011, year, prop_id) %>% 
+  spread(year, prop_id) %>% 
+  rename(prop_id_2004 = `2004`, prop_id_2006 = `2006`) %>% 
+  ggplot(., aes(x = prop_id_2004, y = prop_id_2006)) + 
+  geom_point(alpha = 0.1) + 
+  geom_abline(colour = "blue", slope = 1, intercept = 0) + 
+  stat_smooth(colour = "red", linetype = "dashed") + 
+  labs(
+    x = "ID prop, 2004", y = "ID prop, 2006",
+    title = "2006 against 2004"
+       ) -> s_04_06
+
+simd_2011_reweighted %>% 
+  filter(year %in% c(2006, 2009)) %>% 
+  mutate(prop_id = pop_incomedeprived / pop_total) %>% 
+  select(dz_2011, year, prop_id) %>% 
+  spread(year, prop_id) %>% 
+  rename(prop_id_2006 = `2006`, prop_id_2009 = `2009`) %>% 
+  ggplot(., aes(x = prop_id_2006, y = prop_id_2009)) + 
+  geom_point(alpha = 0.1) + 
+  geom_abline(colour = "blue", slope = 1, intercept = 0) + 
+  stat_smooth(colour = "red", linetype = "dashed") + 
+  labs(
+    x = "ID prop, 2006", y = "ID prop, 2009",
+    title = "2009 against 2006"
+    ) -> s_06_09
+
+simd_2011_reweighted %>% 
+  filter(year %in% c(2009, 2012)) %>% 
+  mutate(prop_id = pop_incomedeprived / pop_total) %>% 
+  select(dz_2011, year, prop_id) %>% 
+  spread(year, prop_id) %>% 
+  rename(prop_id_2009 = `2009`, prop_id_2012 = `2012`) %>% 
+  ggplot(., aes(x = prop_id_2009, y = prop_id_2012)) + 
+  geom_point(alpha = 0.1) + 
+  geom_abline(colour = "blue", slope = 1, intercept = 0) + 
+  stat_smooth(colour = "red", linetype = "dashed") + 
+  labs(
+    x = "ID prop, 2009", y = "ID prop, 2012",
+    title = "2012 against 2009"
+  ) -> s_09_12
+
+plot_grid(s_04_12, s_04_06, s_06_09, s_09_12, nrow = 2)
+
+```
+
+There were indications in the maps above that the greatest changes occurred between the 2009 and 2012 SIMD releases, suggesting that the temporal association between 2004 and 2009 ID proportions would likely be strong compared with the temporal association between 2009 and 2012.
+
+The first of the plots, 2004 compared with 2012, shows strong spatial dependence in areas with low to moderate ID proportions, but much less temporal dependence in areas with high ID proportions in 2004.
+The second subfigure plots 2004 against 2006 values, and shows a near linear temporal dependence for the complete range of 2004 values, with a global trend towards lower ID proportions throughout Scotland.
+The temporal relationship between 2006 and 2009 starts to become nonlinear at the upper range of 2006 values, and this nonlinearity at the upper range becomes much stronger between 2009 and 2012.
+
+These plots indicate a complex change in the temporal dependence of ID proportions, likely occurring over the period 2008 to 2012.
+SIMD 2016 values will be invaluable in indicating whether these changes have continued, altered or settled between 2012 and 2016.
+
 ## Plots of the distance/deprivation relationship
@@
 
-## RCI calculations
-
-This section will present the results of the RCI calculations, i.e. more formal attempt to replicate the results of Kavanagh, Pryce & Lee (2016) using a different indicator and data source.
-
 # Discussion

### 2016-08-04 Script.R

@@ -4,14 +4,21 @@ require(pacman)
 
 pacman::p_load(
   readr, readxl,
-  purrr, stringr, tidyr, dplyr, 
-  broom,
-  rgeos,
-  ggplot2, cowplot,
-  tmap
+  purrr, Rcpp, MCMCpack,
+  stringr, tidyr, dplyr, broom,
+  rgeos, shapefiles, sp, spdep, truncdist,
+  maptools,   tmap,
+  ggplot2, cowplot, RColorBrewer
 )
 
+# My script
 source("scripts/do_dz2011_data_management.R")
+
+# Gavin's scripts
+source('scripts/from_gavin/RCI.R')
+source('scripts/from_gavin/binomial.MCARleroux.R')
+Rcpp::sourceCpp('scripts/from_gavin/aqmen.cpp')
+
 
 # Check this works - plot distance to nearest centre
 # dz_2011 %>% 
@@ -176,6 +183,291 @@ simple_results %>%
   geom_vline(xintercept = 0, linetype = "dashed") + 
   labs(x = "change in total population with distance", y = "change in income deprived population with distance")
 
+
+
+
+
+# Do RCI  -----------------------------------------------------------------
+
+# Tasks 
+
+# 1) Split Scotland up into four cities, dzs within 15km, for 2004 and 2012 
+# 2) produce spatial structure objects 
+
+
+# #### Compute the spatial autocorrelation using Moran's I
+# moran.mc(x=sheffield.map.final@data$prop.eu15.2001, listw=W.list.city, nsim=10000)
+# moran.mc(x=sheffield.map.final@data$prop.eu15.2011, listw=W.list.city, nsim=10000)
+# moran.mc(x=sheffield.map.final@data$prop.eu12.2011, listw=W.list.city, nsim=10000)
+
+#### The temporal dependence
+# plot(prop.eu15.2001, prop.eu15.2011, col="red", pch=19, xlab="2001", ylab="2011")
+# abline(0,1, col="blue")
+# cor.test(prop.eu15.2001,prop.eu15.2011)
+# 
+# simd_2011_reweighted %>% 
+#   filter(year %in% c(2004, 2012)) %>% 
+#   mutate(prop_id = pop_incomedeprived / pop_total) %>% 
+#   select(dz_2011, year, prop_id) %>% 
+#   spread(year, prop_id) %>% 
[…]

### 2016-08-04 scripts/from_gavin/do_rci_code.R

@@ -1,0 +1,347 @@
+# Gavin's RCI code (in other context)
+
+############################### Inference on COB segregation file_ 14/06/2016
+
+
+#### Sheffield RCI ANALYSES
+
+#setwd("C:/Users/Guanpeng/Dropbox/Migrants in the city confernce papers/segregation measures/analysis")
+setwd("/Users/gavin/Dropbox/Migrants in the city confernce papers/segregation measures/analysis")
+
+#### Load the libraries needed
+library(Rcpp)
+library(shapefiles)
+library(sp)
+library(spdep)
+library(CARBayes)
+library(MCMCpack)
+library(truncdist)
+
+library(rgdal)
+library(maptools)
+require(RColorBrewer)
+library(dplyr)
+#### Source the functions for running the model
+source('RCI.R')
+source('binomial.MCARleroux.R')
+Rcpp::sourceCpp('aqmen.cpp')
+
+
+#### Check the boundary changes for different TTWAs. if not much not unchanged LSOAs or DATA zones would be used
+### Import the English COB data in 2001
+cob2001 <- read.csv("Data/England_LSOA_2001.csv",header=TRUE,stringsAsFactors = FALSE)
+head(cob2001)
+str(cob2001)
+summary(cob2001)
+
+### Import the English COB data in 2001
+cob2011 <- read.csv("Data/England_LSOA_2011.csv",header=TRUE,stringsAsFactors = FALSE)
+head(cob2011)
+str(cob2011)
+summary(cob2011)
+
+#### For the three cities (London, Manchester and Sheffield), check the LSOA boundary changes during the two census years
+# link LSOAs to the TTWA defined in 2007
+lsoa2ttwa <- read.csv("Data/LSOA01_TTWA07_UK_LU.csv",header = TRUE, stringsAsFactors = FALSE)
+match.ind <- match(cob2001$code,lsoa2ttwa$LSOA01CD)
+sum(is.na(match.ind))
+cob2001$ttwa <- lsoa2ttwa$TTWA07NM[match.ind]
+data.frame(table(cob2001$ttwa))
+
+# changes of LSOAs boundaries
+changes.boundary <- read.csv("Data/LSOA01_LSOA11_EW_LUv2.csv",header = TRUE,stringsAsFactors = FALSE) %>% tbl_df() %>%
[…]

### 2016-08-05 markdown/draft.RmD
update script, tidying up and automating bayesian model run functions

@@
 2.5% CrI   | 0.142     | 0.091      | -0.054
-97.5% CrI  | 0.146     | 0.095      | -0.048
+97.5% CrI  | 0.146     | 0.096      | -0.047
 
@@
 97.5% CrI  | 0.370   | 0.317    | -0.051
+
+#### Dundee
+
+Quantile   | RCI, 2004 | RCI, 2012  | Change  
+---------- | --------  | ---------- | -------  
+median     | 0.304     | 0.256      | -0.048
+2.5% CrI   | 0.298     | 0.251      | -0.056
+97.5% CrI  | 0.309     | 0.262      | -0.040
+
+Quantile   | D, 2004 | D, 2012  | Change  
+---------- | ------- | -------- | ------- 
+median     | 0.414   | 0.377    | -0.037
+2.5% CrI   | 0.409   | 0.372    | -0.044
+97.5% CrI  | 0.418   | 0.382    | -0.030
+
+#### Aberdeen
+
+Quantile   | RCI, 2004 | RCI, 2012  | Change  
+---------- | --------  | ---------- | -------  
+median     | 0.218     | 0.187      | -0.031
+2.5% CrI   | 0.213     | 0.181      | -0.040
+97.5% CrI  | 0.224     | 0.193      | -0.022
+
+Quantile   | D, 2004 | D, 2012  | Change  
+---------- | ------- | -------- | ------- 
+median     | 0.385   | 0.360    | -0.024
+2.5% CrI   | 0.379   | 0.355    | -0.032
+97.5% CrI  | 0.390   | 0.366    | -0.017
+
+#### Edinburgh
+
+Quantile   | RCI, 2004 | RCI, 2012  | Change  
+---------- | --------  | ---------- | -------  
+median     | 0.013     | -0.013     | -0.025
+2.5% CrI   | 0.009     | -0.016     | -0.030
+97.5% CrI  | 0.016     | -0.009     | -0.020
+
+Quantile   | D, 2004 | D, 2012  | Change  
+---------- | ------- | -------- | ------- 
+median     | 0.385   | 0.319    | -0.067
+2.5% CrI   | 0.382   | 0.316    | -0.071
+97.5% CrI  | 0.389   | 0.322    | -0.062
+
+<!-- > round(apply(indicators_edinburgh, 2, quantile, c(0.5, 0.025, 0.975)),3) -->
+<!--       RCI2004 RCI2012 D2004 D2012 -->
+<!-- 50%     0.013  -0.013 0.385 0.319 -->
+<!-- 2.5%    0.009  -0.016 0.382 0.316 -->
+<!-- 97.5%   0.016  -0.009 0.389 0.322 -->
+<!-- > ## Differences in RCI and D in 2011 - 2001 -->
+<!-- > round(quantile(indicators_edinburgh[ ,2] - indicators_edinburgh[ ,1], c(0.5, 0.025, 0.975)),3) -->
+<!--    50%   2.5%  97.5%  -->
+<!-- -0.025 -0.030 -0.020  -->
+<!-- > round(quantile(indicators_edinburgh[ ,4] - indicators_edinburgh[ ,3], c(0.5, 0.025, 0.975)),3) -->
+<!--    50%   2.5%  97.5%  -->
+<!-- -0.067 -0.071 -0.062  -->
+<!-- >  -->
+<!-- > indicators_aberdeen <- do_model("Aberdeen") -->
+<!-- Keys match perfectly. -->
+
+<!-- Keys match perfectly. -->
+
+<!-- Joining by: "dz_2011" -->
+<!-- data key contains NA's, which are ignored -->
+<!-- Over coverage. 6628 out of 6975 unmatched data records: S01006789, S01006790, S01006791, S01006792, S01006793, ... -->
+<!-- Setting up the model -->
+<!-- Collecting 20000 samples -->
+<!--   |====================================================================================| 100% -->
+<!-- Summarising results -->
+<!--  finished in  133.3 seconds -->
+<!-- > round(apply(indicators_aberdeen, 2, quantile, c(0.5, 0.025, 0.975)),3) -->
+<!--       RCI2004 RCI2012 D2004 D2012 -->
+<!-- 50%     0.218   0.187 0.385 0.360 -->
+<!-- 2.5%    0.213   0.181 0.379 0.355 -->
+<!-- 97.5%   0.224   0.193 0.390 0.366 -->
+<!-- > ## Differences in RCI and D in 2011 - 2001 -->
+<!-- > round(quantile(indicators_aberdeen[ ,2] - indicators_aberdeen[ ,1], c(0.5, 0.025, 0.975)),3) -->
+<!--    50%   2.5%  97.5%  -->
+<!-- -0.031 -0.040 -0.022  -->
+<!-- > round(quantile(indicators_aberdeen[ ,4] - indicators_aberdeen[ ,3], c(0.5, 0.025, 0.975)),3) -->
+<!--    50%   2.5%  97.5%  -->
+<!-- -0.024 -0.032 -0.017  -->
 
@@
 Within these plots the blue line indicates parity between values on the x and y axis, and the red dashed line shows a nonlinear smoother function between the two ID proportions being plotted, and hence the level of deviation between the actual relationship over the two years and parity.
+
 ```{r show_temporal_dep, echo = FALSE, warning = FALSE, EVAL = TRUE}
@@
 (The smoother line at distances less than 1km should likely be ignored for most cities as it involves extrapolation based on very few observations).
+
 ```{r plot_changetotalpop, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
@@
 ```
+
 This plot suggests that there have been substantial increases in population sizes in Aberdeen, Edimburgh and Glasgow over this period, in datazones closer to the city centre, with the greatest changes occurring near the centre of Glasgow.
@@
 output areas) to recalculate and reapportion SIMD scores to 2011 datazone geographies.
-As with all analyses, the additional benefit of alternative and more sophisticated approaches should be weighted against the additional time cost involved, and a judgement made about whether the ratio of marginal benefits to marginal costs represents a cost effective use of researcher time.
 

### 2016-08-05 Script.R

@@ -278,8 +278,10 @@ do_model <- function(place, initial_dist = 18000){
         shp = . , data = centre_distance, 
         key.shp = "DataZone", key.data = "datazone"
       ) %>% 
-      .[.$nearest_centre == "Glasgow",] %>% 
+      .[.$nearest_centre == place,] %>% 
       .[.$distance_to_centre <= this_dist,] -> dz_city
+    # plot(dz_city, main = this_dist)
+    # browser()
 
     w <-  poly2nb(dz_city)   
     if (any(card(w) == 0)){
@@ -316,7 +318,7 @@ do_model <- function(place, initial_dist = 18000){
     .[!is.na(.$pop_total_2004),] -> dz_city
   
   
-  w_nb <- poly2nb(dz_city)
+  W_nb <- poly2nb(dz_city)
   # distance to allow the 'islands' to be included 
   W_list <- nb2listw(W_nb, style = "B")
   W <- nb2mat(W_nb, style = "B")
@@ -355,6 +357,7 @@ do_model <- function(place, initial_dist = 18000){
   
   all_2004 <- as.integer(N.mat[,1])
   all_2012 <- as.integer(N.mat[,2])
+  
   for(i in 1:n.keep)
   {
     ## Compute the probability and fitted values for the ith posterior sample
@@ -364,8 +367,8 @@ do_model <- function(place, initial_dist = 18000){
     fitted.mat <- N.mat * prob.mat
     
     ## Compute the RCI for both years
-    indicators.post[i, 1] <-RCI(fitted.mat[ ,1], as.integer(N.mat[,1]), dist.order)
-    indicators.post[i, 2] <-RCI(fitted.mat[ ,2], as.integer(N.mat[,2]), dist.order)
+    indicators.post[i, 1] <- RCI(fitted.mat[ ,1], as.integer(N.mat[,1]), dist.order)
+    indicators.post[i, 2] <- RCI(fitted.mat[ ,2], as.integer(N.mat[,2]), dist.order)
     
     ## Compute D for both years
     p_2004 <- prob.mat[ ,1]
@@ -380,24 +383,258 @@ do_model <- function(place, initial_dist = 18000){
   indicators.post  
 }
 
-
-indicators_dundee <- do_model("Dundee")
 indicators_aberdeen <- do_model("Aberdeen")
+indicators_dundee <- do_model("Dundee")
[…]

### 2016-08-05 markdown/draft.RmD
do and tidy running and outputs of models

@@
 
-<!-- > round(apply(indicators_edinburgh, 2, quantile, c(0.5, 0.025, 0.975)),3) -->
-<!--       RCI2004 RCI2012 D2004 D2012 -->
-<!-- 50%     0.013  -0.013 0.385 0.319 -->
-<!-- 2.5%    0.009  -0.016 0.382 0.316 -->
-<!-- 97.5%   0.016  -0.009 0.389 0.322 -->
-<!-- > ## Differences in RCI and D in 2011 - 2001 -->
-<!-- > round(quantile(indicators_edinburgh[ ,2] - indicators_edinburgh[ ,1], c(0.5, 0.025, 0.975)),3) -->
-<!--    50%   2.5%  97.5%  -->
-<!-- -0.025 -0.030 -0.020  -->
-<!-- > round(quantile(indicators_edinburgh[ ,4] - indicators_edinburgh[ ,3], c(0.5, 0.025, 0.975)),3) -->
-<!--    50%   2.5%  97.5%  -->
-<!-- -0.067 -0.071 -0.062  -->
-<!-- >  -->
-<!-- > indicators_aberdeen <- do_model("Aberdeen") -->
-<!-- Keys match perfectly. -->
+### Change in RCI and D over time
 
-<!-- Keys match perfectly. -->
+```{r grab_posteriors, eval = TRUE, cache = FALSE, warning = FALSE, echo = FALSE}
 
-<!-- Joining by: "dz_2011" -->
-<!-- data key contains NA's, which are ignored -->
-<!-- Over coverage. 6628 out of 6975 unmatched data records: S01006789, S01006790, S01006791, S01006792, S01006793, ... -->
-<!-- Setting up the model -->
-<!-- Collecting 20000 samples -->
-<!--   |====================================================================================| 100% -->
-<!-- Summarising results -->
-<!--  finished in  133.3 seconds -->
-<!-- > round(apply(indicators_aberdeen, 2, quantile, c(0.5, 0.025, 0.975)),3) -->
-<!--       RCI2004 RCI2012 D2004 D2012 -->
-<!-- 50%     0.218   0.187 0.385 0.360 -->
-<!-- 2.5%    0.213   0.181 0.379 0.355 -->
-<!-- 97.5%   0.224   0.193 0.390 0.366 -->
-<!-- > ## Differences in RCI and D in 2011 - 2001 -->
-<!-- > round(quantile(indicators_aberdeen[ ,2] - indicators_aberdeen[ ,1], c(0.5, 0.025, 0.975)),3) -->
-<!--    50%   2.5%  97.5%  -->
-<!-- -0.031 -0.040 -0.022  -->
-<!-- > round(quantile(indicators_aberdeen[ ,4] - indicators_aberdeen[ ,3], c(0.5, 0.025, 0.975)),3) -->
-<!--    50%   2.5%  97.5%  -->
-<!-- -0.024 -0.032 -0.017  -->
+all_posteriors <- read_csv("data/all_posterior_draws.csv") # Results from running the many models
+
+all_posteriors  %>% 
+  group_by(place, measure, period)  %>% 
+  summarise(
+    lower = quantile(value, 0.025), 
+    median = quantile(value, 0.500), 
+    upper = quantile(value, 0.975)
+            ) -> posterior_summaries
+```
+
+Unlike the census-based variables used in Kavanagh, Lee & Pryce (2016), there are SIMD releases for 2004, 2006, 2009, 2012, and 2016, meaning in RCI and D scores within the period 2001 and 2011 can be observed.
+The Bayesian spatial modelling approach used by Kavanagh, Lee & Pryce (2016) was applied to each of the SIMD releases, and median RCI and D estimates, along with 95% CrIs, were produced for each release and for each city.
+
+The graph of RCI over time is shown below.
+
+```{r plot_change_rci, eval = TRUE, echo = FALSE, warning = FALSE}
+# Do RCI graph
+posterior_summaries %>% 
+  filter(measure == "RCI") %>% 
+   mutate(period = as.factor(period)) %>% 
+   ggplot(., 
+          aes(
+            x = period, group = place, 
+            y = median, colour = place,
+            shape = place)) + 
+   geom_line() + geom_point() + 
+   scale_y_continuous(limits = c(0, 0.4) ) +
+   geom_line(aes(y = upper), linetype = "dashed") + 
+   geom_line(aes(y = lower), linetype = "dashed") + 
+   labs(
+     y = "RCI of income deprived", x = "Year", 
+     title = "Change in RCI", 
+     colour = "City", shape = "City") +
+   theme(legend.position = c(0.8, 0.8))
+
+```
+
+This shows that the greatest falls in RCI occurred between the 2006 and 2009 SIMD releases
+
+The plot of changes in D over time is shown below
+
+```{r plot_change_d, eval = TRUE, echo = FALSE, warning = FALSE}
+# Do D graph
+
+posterior_summaries %>% 
+  filter(measure == "D") %>% 
+  mutate(period = as.factor(period)) %>% 
+  ggplot(., 
+         aes(
+           x = period, group = place, 
+           y = median, colour = place,
+           shape = place)) + 
+  geom_line() + geom_point() + 
+  scale_y_continuous(limits = c(0, 0.5) ) +
+  geom_line(aes(y = upper), linetype = "dashed") + 
+  geom_line(aes(y = lower), linetype = "dashed") + 
+  labs(
+    y = "D of income deprived", x = "Year", 
+    title = "Change in D", 
+    colour = "City", shape = "City") +
+  theme(legend.position = c(0.8, 0.8))
+
+```
+
+Similarly, the the greatest fall in D occurred between 2006 and 2009.
 

### 2016-08-05 Script.R

@@ -254,573 +254,51 @@ simple_results %>%
 # plot_grid(s_04_12, s_04_09, s_09_12, nrow = 1)
 # 
 
+all_posteriors <- read_csv("data/all_posterior_draws.csv") # Results from running the many models
 
+all_posteriors  %>% 
+  group_by(place, measure, period)  %>% 
+  summarise(
+    lower = quantile(value, 0.025), 
+    median = quantile(value, 0.500), 
+    upper = quantile(value, 0.975)
+            ) -> posterior_summaries
 
-################################
-#### Fit the model
-################################
-#### MCMC quantities
-burnin <- 10000
-n.sample <- 20000
-thin <- 10
-n.keep <- (n.sample - burnin)/thin
-
-# Fit the model for Glasgow first 
-
-do_model <- function(place, initial_dist = 18000){
-  
-  this_dist <- initial_dist
-  has_islands <- T
-  
-  while(has_islands){
-    dz_2011  %>% 
-      append_data(
-        shp = . , data = centre_distance, 
-        key.shp = "DataZone", key.data = "datazone"
-      ) %>% 
-      .[.$nearest_centre == place,] %>% 
-      .[.$distance_to_centre <= this_dist,] -> dz_city
-    # plot(dz_city, main = this_dist)
-    # browser()
-
-    w <-  poly2nb(dz_city)   
-    if (any(card(w) == 0)){
-        this_dist <- this_dist + 1000
-    } else {
-      has_islands <- F
-    }
-  }
-  
-  simd_2011_reweighted %>% 
-    filter(year %in% c(2004, 2012)) %>% 
-    select(dz_2011, year, pop_total, pop_incomedeprived) -> tmp
-  
-  tmp %>% 
-    select(-pop_incomedeprived) %>% 
-    mutate(year = paste0("pop_total_", year)) %>% 
-    spread(year, pop_total) -> pops
-  
-  tmp %>% 
-    select(-pop_total) %>% 
-    mutate(year = paste0("pop_id_", year)) %>% 
-    spread(year, pop_incomedeprived) -> incdeps
-  
-  popinc <- pops %>% inner_join(incdeps)
-  rm(tmp, pops, incdeps)
-  
-  
-  dz_city %>% 
-    append_data(
-      shp = ., data = popinc,
[…]

### 2016-08-05 scripts/from_gavin/do_rci_and_d_for_all_permutations.R

@@ -1,0 +1,180 @@
+
+
+################################
+#### Fit the model
+################################
+#### MCMC quantities
+burnin <- 10000
+n.sample <- 20000
+thin <- 10
+n.keep <- (n.sample - burnin)/thin
+
+
+# Now to revise do_model to run all available years 
+
+do_model <- function(place, initial_dist = 18000){
+  
+  this_dist <- initial_dist
+  has_islands <- T
+  
+  while(has_islands){
+    dz_2011  %>% 
+      append_data(
+        shp = . , data = centre_distance, 
+        key.shp = "DataZone", key.data = "datazone"
+      ) %>% 
+      .[.$nearest_centre == place,] %>% 
+      .[.$distance_to_centre <= this_dist,] -> dz_city
+    # plot(dz_city, main = this_dist)
+    # browser()
+    
+    w <-  poly2nb(dz_city)   
+    if (any(card(w) == 0)){
+      this_dist <- this_dist + 1000
+    } else {
+      has_islands <- F
+    }
+  }
+  
+  simd_2011_reweighted %>% 
+    select(dz_2011, year, pop_total, pop_incomedeprived) -> tmp
+  
+  tmp %>% 
+    select(-pop_incomedeprived) %>% 
+    mutate(year = paste0("pop_total_", year)) %>% 
+    spread(year, pop_total) -> pops
+  
+  tmp %>% 
+    select(-pop_total) %>% 
+    mutate(year = paste0("pop_id_", year)) %>% 
+    spread(year, pop_incomedeprived) -> incdeps
+  
+  popinc <- pops %>% inner_join(incdeps)
+  rm(tmp, pops, incdeps)
+  
+  
+  dz_city %>% 
+    append_data(
+      shp = ., data = popinc,
+      key.shp = "DataZone", key.data = "dz_2011",
+      ignore.na = T
+    ) %>% 
+    .[!is.na(.$pop_total_2004),] -> dz_city
+  
+  
+  W_nb <- poly2nb(dz_city)
+  # distance to allow the 'islands' to be included 
+  W_list <- nb2listw(W_nb, style = "B")
+  W <- nb2mat(W_nb, style = "B")
+  n <- nrow(W)
+  
+  # Need to remove unconnected datazones ('islands')
+  
+  
+  
[…]

### 2016-08-05 markdown/draft.RmD
update markdown doc - update date; add tables direct from values

@@
 author: "Jon Minton"
-date: "4 August 2016"
+date: "5 August 2016"
 output: html_document
@@
 
+```
+
+```{r grab_posteriors, eval = TRUE, cache = FALSE, warning = FALSE, echo = FALSE}
+
+all_posteriors <- read_csv("data/all_posterior_draws.csv") # Results from running the many models
+
+all_posteriors  %>% 
+  group_by(place, measure, period)  %>% 
+  summarise(
+    lower = quantile(value, 0.025), 
+    median = quantile(value, 0.500), 
+    upper = quantile(value, 0.975)
+            ) -> posterior_summaries
 ```
@@
 ```{r show_map, eval = TRUE, echo = FALSE, cache = FALSE, message = FALSE, warning = FALSE}
+
 dz_2011 %>%
@@
 
-For Glasgow, the RCI scores (with 95% credible intervals, CrIs) for 2004 and 2012, and for the difference between these two periods, are as follows:
+### RCI, Aberdeen
 
-Quantile   | RCI, 2004 | RCI, 2012  | Change  
----------- | --------  | ---------- | -------  
-median     | 0.144     | 0.093      | -0.051
-2.5% CrI   | 0.142     | 0.091      | -0.054
-97.5% CrI  | 0.146     | 0.096      | -0.047
+```{r, eval = TRUE, echo = FALSE, warning = FALSE, results = 'asis'}
+posterior_summaries %>% 
+  ungroup() %>% 
+  filter(place == "Aberdeen", measure == "RCI") %>% 
+  gather(key = "bound", value = "value", lower:upper) %>% 
+  select(period, bound, value) %>% 
+  mutate(value = round(value, 3)) %>% 
+  spread(bound, value) %>% 
+  select(period, median, lower, upper) %>% 
+  kable(., format = "markdown", caption = "RCI, Aberdeen")
 
-From this we can conclude that there has been a statistically significant fall in RCI over these two periods using the SIMD income deprivation measure, as there had been for out-of-work benefits-based measures for 2001 and 2011 in Kavanagh at al 2016.
+```
 
-As with Kavanagh et al 2016 Dissimilarity index scores (D) were also calculated for both periods, along with change between them.
-These are shown in the table below, and also show statistically significant falls.
+### RCI, Dundee
 
-Quantile   | D, 2004 | D, 2012  | Change  
----------- | ------- | -------- | ------- 
-median     | 0.368   | 0.315    | -0.053
-2.5% CrI   | 0.366   | 0.313    | -0.056
-97.5% CrI  | 0.370   | 0.317    | -0.051
+```{r, eval = TRUE, echo = FALSE, warning = FALSE, results = 'asis'}
+posterior_summaries %>% 
+  ungroup() %>% 
+  filter(place == "Dundee", measure == "RCI") %>% 
+  gather(key = "bound", value = "value", lower:upper) %>% 
+  select(period, bound, value) %>% 
+  mutate(value = round(value, 3)) %>% 
+  spread(bound, value) %>% 
+  select(period, median, lower, upper) %>% 
+  kable(., format = "markdown", caption = "RCI, Dundee")
 
-#### Dundee
+```
 
-Quantile   | RCI, 2004 | RCI, 2012  | Change  
----------- | --------  | ---------- | -------  
-median     | 0.304     | 0.256      | -0.048
-2.5% CrI   | 0.298     | 0.251      | -0.056
-97.5% CrI  | 0.309     | 0.262      | -0.040
+### RCI, Edinburgh
 
-Quantile   | D, 2004 | D, 2012  | Change  
----------- | ------- | -------- | ------- 
-median     | 0.414   | 0.377    | -0.037
-2.5% CrI   | 0.409   | 0.372    | -0.044
-97.5% CrI  | 0.418   | 0.382    | -0.030
+```{r, eval = TRUE, echo = FALSE, warning = FALSE, results = 'asis'}
+posterior_summaries %>% 
+  ungroup() %>% 
+  filter(place == "Edinburgh", measure == "RCI") %>% 
+  gather(key = "bound", value = "value", lower:upper) %>% 
+  select(period, bound, value) %>% 
+  mutate(value = round(value, 3)) %>% 
+  spread(bound, value) %>% 
+  select(period, median, lower, upper) %>% 
+  kable(., format = "markdown", caption = "RCI, Edinburgh")
 
-#### Aberdeen
+```
 
-Quantile   | RCI, 2004 | RCI, 2012  | Change  
----------- | --------  | ---------- | -------  
-median     | 0.218     | 0.187      | -0.031
-2.5% CrI   | 0.213     | 0.181      | -0.040
-97.5% CrI  | 0.224     | 0.193      | -0.022
+### RCI, Glasgow
 
-Quantile   | D, 2004 | D, 2012  | Change  
----------- | ------- | -------- | ------- 
-median     | 0.385   | 0.360    | -0.024
-2.5% CrI   | 0.379   | 0.355    | -0.032
-97.5% CrI  | 0.390   | 0.366    | -0.017
+```{r, eval = TRUE, echo = FALSE, warning = FALSE, results = 'asis'}
+posterior_summaries %>% 
+  ungroup() %>% 
+  filter(place == "Glasgow", measure == "RCI") %>% 
+  gather(key = "bound", value = "value", lower:upper) %>% 
+  select(period, bound, value) %>% 
+  mutate(value = round(value, 3)) %>% 
+  spread(bound, value) %>% 
+  select(period, median, lower, upper) %>% 
+  kable(., format = "markdown", caption = "RCI, Glasgow")
 
-#### Edinburgh
+```
 
-Quantile   | RCI, 2004 | RCI, 2012  | Change  
----------- | --------  | ---------- | -------  
-median     | 0.013     | -0.013     | -0.025
-2.5% CrI   | 0.009     | -0.016     | -0.030
-97.5% CrI  | 0.016     | -0.009     | -0.020
+## Dissimilarity calculations
 
-Quantile   | D, 2004 | D, 2012  | Change  
----------- | ------- | -------- | ------- 
-median     | 0.385   | 0.319    | -0.067
-2.5% CrI   | 0.382   | 0.316    | -0.071
-97.5% CrI  | 0.389   | 0.322    | -0.062
+This section will present the results of the Dissimilarity index calculations, using the Bayesian modelling approach used in Kavanagh et al 2016.
+
+### D, Aberdeen
+
+```{r, eval = TRUE, echo = FALSE, warning = FALSE, results = 'asis'}
+posterior_summaries %>% 
+  ungroup() %>% 
+  filter(place == "Aberdeen", measure == "D") %>% 
+  gather(key = "bound", value = "value", lower:upper) %>% 
+  select(period, bound, value) %>% 
+  mutate(value = round(value, 3)) %>% 
+  spread(bound, value) %>% 
+  select(period, median, lower, upper) %>% 
+  kable(., format = "markdown", caption = "D, Aberdeen")
+
+```
+
+### D, Dundee
+
+```{r, eval = TRUE, echo = FALSE, warning = FALSE, results = 'asis'}
+posterior_summaries %>% 
+  ungroup() %>% 
+  filter(place == "Dundee", measure == "D") %>% 
+  gather(key = "bound", value = "value", lower:upper) %>% 
+  select(period, bound, value) %>% 
+  mutate(value = round(value, 3)) %>% 
+  spread(bound, value) %>% 
+  select(period, median, lower, upper) %>% 
+  kable(., format = "markdown", caption = "D, Dundee")
+
+```
+
+### D, Edinburgh
+
+```{r, eval = TRUE, echo = FALSE, warning = FALSE, results = 'asis'}
+posterior_summaries %>% 
+  ungroup() %>% 
+  filter(place == "Edinburgh", measure == "D") %>% 
+  gather(key = "bound", value = "value", lower:upper) %>% 
+  select(period, bound, value) %>% 
+  mutate(value = round(value, 3)) %>% 
+  spread(bound, value) %>% 
+  select(period, median, lower, upper) %>% 
+  kable(., format = "markdown", caption = "D, Edinburgh")
+
+```
+
+### RCI, Glasgow
+
+```{r, eval = TRUE, echo = FALSE, warning = FALSE, results = 'asis'}
+posterior_summaries %>% 
+  ungroup() %>% 
+  filter(place == "Glasgow", measure == "D") %>% 
+  gather(key = "bound", value = "value", lower:upper) %>% 
+  select(period, bound, value) %>% 
+  mutate(value = round(value, 3)) %>% 
+  spread(bound, value) %>% 
+  select(period, median, lower, upper) %>% 
+  kable(., format = "markdown", caption = "D, Glasgow")
+
+```
 
 ### Change in RCI and D over time
-
-```{r grab_posteriors, eval = TRUE, cache = FALSE, warning = FALSE, echo = FALSE}
-
-all_posteriors <- read_csv("data/all_posterior_draws.csv") # Results from running the many models
-
-all_posteriors  %>% 
-  group_by(place, measure, period)  %>% 
-  summarise(
-    lower = quantile(value, 0.025), 
-    median = quantile(value, 0.500), 
-    upper = quantile(value, 0.975)
-            ) -> posterior_summaries
-```
 

### 2016-08-16 Script.R
redid carbayes code

@@ -5,6 +5,7 @@ require(pacman)
 pacman::p_load(
   readr, readxl,
   purrr, Rcpp, MCMCpack,
+  CARBayes,
   stringr, tidyr, dplyr, broom,
   rgeos, shapefiles, sp, spdep, truncdist,
   maptools,   tmap,

### 2016-08-16 scripts/from_gavin/do_rci_and_d_for_all_permutations.R

@@ -74,27 +74,36 @@ do_model <- function(place, initial_dist = 18000){
   
   #### Format the data
   Y.mat <- cbind(
-    dz_city@data$pop_id_2004,
-    dz_city@data$pop_id_2006, 
-    dz_city@data$pop_id_2009, 
-    dz_city@data$pop_id_2012 
+    as.integer(dz_city@data$pop_id_2004),
+    as.integer(dz_city@data$pop_id_2006), 
+    as.integer(dz_city@data$pop_id_2009), 
+    as.integer(dz_city@data$pop_id_2012) 
   )
-  Y <- as.integer(t(Y.mat)) # Noninteger values as reweighted
   
   N.mat <- cbind(
-    dz_city@data$pop_total_2004,
-    dz_city@data$pop_total_2006, 
-    dz_city@data$pop_total_2009, 
-    dz_city@data$pop_total_2012 
+    as.integer(dz_city@data$pop_total_2004),
+    as.integer(dz_city@data$pop_total_2006), 
+    as.integer(dz_city@data$pop_total_2009), 
+    as.integer(dz_city@data$pop_total_2012) 
   )
-  N <- as.integer(t(N.mat)) # Noninteger values as reweighted
+
+
   
   #### Run the model
-  model <- binomial.MCARleroux(
-    formula=Y~1, trials=N, W=W, 
-    burnin=burnin, n.sample=n.sample, thin=thin
-  )
-  model$summary.results
+  Y <- Y.mat[,1] ; N <- N.mat[,1]
+  model_01 <- S.CARleroux(Y ~ 1, trials = N, W = W, family = "binomial", burnin = burnin, n.sample = n.sample, thin = thin )  
+
+  Y <- Y.mat[,2] ; N <- N.mat[,2]
+  model_02 <- S.CARleroux(Y ~ 1, trials = N, W = W, family = "binomial", burnin = burnin, n.sample = n.sample, thin = thin )  
+
+  Y <- Y.mat[,3] ; N <- N.mat[,3]
+  model_03 <- S.CARleroux(Y ~ 1, trials = N, W = W, family = "binomial", burnin = burnin, n.sample = n.sample, thin = thin )  
+
+  Y <- Y.mat[,4] ; N <- N.mat[,4]
[…]

### 2016-08-17 markdown/draft.RmD
make tweaks to the markdown doc

@@
 The graph of RCI over time is shown below.
+The density of RCI estimates for each place and in each year is shown using a violin plot, with longer shapes indicating a greater range of posterior estimates.
+The median RCI scores are indicated as points, linked for each city with dashed lines.
 
@@
 # Do RCI graph
-posterior_summaries %>% 
+
+all_posteriors  %>% 
   filter(measure == "RCI")   %>% 
-   mutate(period = as.factor(period)) %>% 
-   ggplot(., 
-          aes(
-            x = period, group = place, 
-            y = median, colour = place,
-            shape = place)) + 
-   geom_line() + geom_point() + 
-   scale_y_continuous(limits = c(0, 0.4) ) +
-   geom_line(aes(y = upper), linetype = "dashed") + 
-   geom_line(aes(y = lower), linetype = "dashed") + 
+  group_by(place, period)  %>% 
+  mutate(med_val = median(value))  %>%  
+  mutate(pp = paste0(place, period))  %>% 
+  ggplot(.) + 
+  geom_violin(
+    aes(x = factor(period), y = value, group = pp, fill = place), 
+    alpha = 0.2, position = "identity"
+  ) + 
+  geom_line(
+    aes(x = factor(period), y = med_val, group = place, colour = place), 
+    size = 1.1, linetype = "dashed"
+  ) + 
+  geom_point(
+    aes(x = factor(period), y = med_val, shape = place, colour = place), 
+    size = 3.5
+  ) + 
   labs(
-     y = "RCI of income deprived", x = "Year", 
-     title = "Change in RCI", 
-     colour = "City", shape = "City") +
-   theme(legend.position = c(0.8, 0.8))
+    x = "Year", y = "RCI", colour = "City", 
+    shape = "City", fill = "City", title = "Change in RCI scores over time"
+  ) + 
+  geom_hline(aes(yintercept = 0))
 
@@
 
-This shows that the greatest falls in RCI occurred between the 2006 and 2009 SIMD releases
+The overall trend in each place is for RCI to reduce from 2004 to 2012.
+However within each period the ordering of RCIs between places has not changed, with Dundee showing the most centralisation of poverty, then Aberdeen, and Glasgow, with Edinburgh the most decentralised.
+In terms of trends over time, Glasgow and Dundee both show steadily declining RCI between SIMD releases, whereas for Aberdeen and Edinbugh RCI values increased from 2004 to 2006, before declining in 2009 and 2012.
+From 2009 onwards Edinburgh has negative RCI scores, suggesting a peripheralisation rather than centralisation of poverty.
 
-The plot of changes in D over time is shown below
+The equivalent graph for the Dissimilarity index is shown below:
 
@@
 
-posterior_summaries %>% 
+all_posteriors %>% 
   filter(measure == "D")   %>% 
-  mutate(period = as.factor(period)) %>% 
-  ggplot(., 
-         aes(
-           x = period, group = place, 
-           y = median, colour = place,
-           shape = place)) + 
-  geom_line() + geom_point() + 
-  scale_y_continuous(limits = c(0, 0.5) ) +
-  geom_line(aes(y = upper), linetype = "dashed") + 
-  geom_line(aes(y = lower), linetype = "dashed") + 
+  group_by(place, period)  %>% 
+  mutate(med_val = median(value))  %>%  
+  mutate(pp = paste0(place, period))  %>% 
+  ggplot(.) + 
+  geom_violin(
+    aes(x = factor(period), y = value, group = pp, fill = place), 
+    alpha = 0.2, position = "identity"
+  ) + 
+  geom_line(
+    aes(x = factor(period), y = med_val, group = place, colour = place), 
+    size = 1.1, linetype = "dashed"
+  ) + 
+  geom_point(
+    aes(x = factor(period), y = med_val, shape = place, colour = place), 
+    size = 3.5
+  ) + 
   labs(
-    y = "D of income deprived", x = "Year", 
-    title = "Change in D", 
-    colour = "City", shape = "City") +
-  theme(legend.position = c(0.8, 0.8))
+    x = "Year", y = "D", colour = "City", shape = "City", 
+    fill = "City", title = "Change in D scores over time")
 
@@
 
+As with RCI, Dissimilarity scores have also fallen over time for all four cities.
+The greatest contrast in the trends is between Aberdeen and Edinburgh.
+In 2004 both cities had similar and overlapping D scores, but between 2004 and 2009 D fell rapidly for Edinburgh, while remaining fairly stable in Aberdeen, before falling from 2009 to 2012.
+Glasgow has always had the lowest D scores of the four cities, and Dundee the highest, but since 2009 the distribution of D estimates for Edinburgh have started to overlap those of Glasgow as the rate of decline in D has been slightly greater in Edinburgh than Glasgow.
 Similarly, the the greatest fall in D occurred between 2006 and 2009.
@@
 
+## Table of average proportion income deprived by year
+
+Below is a table showing the percentage if the population in each city (within 15 km of the centre) defined as 'income deprived' according to SIMD.
+In 2004 these percentages ranged from 9.8% in Aberdeen to 20.8% in Glasgow.
+By 2012 they had fallen to 7.9% in Aberdeen to 17.1% om Glasgow.
+In every city except Aberdeen the percentage increased from 2006 to 2009, as would be expected given the onset of the 2008 recession within this period, but then fell again from the 2009 to the 2012 SIMD release.
+
+```{r id_prop_by_year, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+  simd_2011_reweighted  %>% 
+    ungroup() %>%  
+    left_join(centre_distance, by = c("dz_2011" = "datazone")) %>%
+    filter(distance_to_centre < 15000) %>% 
+    group_by(nearest_centre, year) %>% 
+    summarise(pop_total = sum(pop_total), pop_incomedeprived = sum(pop_incomedeprived)) %>% 
+    mutate(prop_id = pop_incomedeprived / pop_total) %>% 
+    mutate(prop_id = 100 * round(prop_id, 3)) %>% 
+    select(City = nearest_centre, prop_id, Year = year) %>% 
+    spread(Year, prop_id) %>% 
+    ungroup() %>% 
+    arrange(`2004`) %>% 
+    kable(., format = "markdown", caption = "Income deprived % by City and Year")
+
+```
+
 ## Plots of the distance/deprivation relationship

### 2016-08-24 markdown/draft.RmD
update code to have placeholders for 2015 SIMD. Update markdown with new figures

@@
 ---
-title: "Decentralisation of Poverty - Replication Analysis"
-author: "Jon Minton"
-date: "5 August 2016"
+title: "The Suburbanisation of Poverty in Scotland’s Cities: Urban Regeneration or Social Cleansing?"
+author: "Jon Minton & Nick Bailey"
+date: "31 August 2016"
 output: html_document
@@
 
-* Work very much in progress
-* Please don't steal
+This is work in progress, not a polished report, but shared on line in order to promote more rapid dissemination and discussion. Comments and constructive criticism welcomed.
 
-# Introduction
+# Acknowledgements
 
-This is the introduction section to the report.
+AQMeN and ESRC funding.
 
-## Background
+# Summary
 
-In 2014 Kavanagh, Lee & Pryce produced the AQMeN briefing report [Poverty in Suburbia: Has Glasgow gone the way of American cities?](http://www.research.ed.ac.uk/portal/files/18805757/RB5_poverty_suburbia.pdf).
-This paper argued that there has been a decentralisation of poverty in Glasgow between 2001 and 2011, and combined a Bayesian spatial statistics approach with the Relative Centralisation Index (RCI) in order to argue firstly that the spatial distribution of poverty has become less decentralised between the two periods, and secondly that the changes in RCE scores are statistically significant.
-The briefing paper defined 'poverty' in three ways, each relating to specific UK state benefit receipts for persons of working age:
+* [CHECK] In all four of Scotland’s major cities, the number of poor or ‘income deprived’ people in inner city areas fell between 2012 and 2016 according to the latest Scottish Index of Multiple Deprivation data, while the number of non-poor people in these areas rose.
+*	[CHECK] In less central areas, the population of both groups was little changed. Poverty therefore became more ‘suburbanised’ in this period.
+*	[CHECK] This continues the trend observed over the last four waves of the SIMD, from 2004 to 2012, again across all four cities. 
+*	[CHECK] The total population in the central areas rose in all four cities between 2004 and 2016.
+This is generally to be welcomed from an environmental point of view.
+It suggests that planning policies have succeeded in encouraging people to live in higher density locations which tend to be better served by public transport.
+*	[CHECK] However the displacement of lower income groups suggests these policies will fail to deliver the socially mixed communities which planners have so frequently advocated as part of their vision for sustainable cities.
+*	[CHECK] The main reason for this change is likely to be the continued reductions in housing subsidies for lower income households.
+The last 20 years have seen the loss of social rented housing and greater reliance on private renting, combined with the reductions in housing benefits.
+As a result, inner city locations appear to be increasingly beyond the reach of poorer households.
 
-1. The proportion of local populations receiving Income Support (IS)
-2. The proportion of local populations receiving Incapacity Benefit (IB)
-3. The proportion of local populations receiving Jobseekers' Allowance (JSA)
+```{r id_props, eval = TRUE, echo = FALSE, cache = FALSE, message = FALSE, warning = FALSE}
+simd_2011_reweighted  %>% 
+    ungroup() %>%  
+    left_join(centre_distance, by = c("dz_2011" = "datazone")) %>% 
+    distinct() %>%
+    rename(City = nearest_centre) %>% 
+    filter(City %in% c("Aberdeen", "Dundee", "Edinburgh", "Glasgow")) %>% 
+    mutate(centralness = cut(
+      distance_to_centre, 
+      breaks = c(0, 8000, 20000, Inf), 
+      labels = c("centre", "periphery", "exurb"))
+    ) %>% 
+  filter(centralness != "exurb") %>% 
+  group_by(City, year, centralness) %>%
+  mutate(pop_nonid = pop_total - pop_incomedeprived) %>% 
+  summarise(pop_nonid = sum(pop_nonid), pop_incomedeprived = sum(pop_incomedeprived)) %>% 
+  summarise(
+    `Income Deprived` = pop_incomedeprived[centralness == "centre"] / sum(pop_incomedeprived),
+    `Not Income Deprived` = pop_nonid[centralness == "centre"] / sum(pop_nonid)
+  ) %>% 
+  gather(key = "type", value = "prop", `Income Deprived`:`Not Income Deprived`) -> tmp
 
-These analyses were expanded to the four major Scottish cities - Aberdeen, Dundee, Edinburgh and Glasgow - for a paper to appear in *Annals of the American Association of Geographers*.
-This paper calculated 95% credible intervals (CrIs) for the difference in RCI from 2001 to 2011 for each of the three measures, and each of the four cities, between 2001 and 2011.
-It found falls in RCI for each benefit type, and for all four cities, which were statistically significant.
+tmp %>% filter(type == "Income Deprived") %>% 
+  ggplot(., 
+         aes(
+            x = year, y = prop 
+            )
+         ) + 
+  geom_line() + geom_point() +
+  facet_wrap( ~ City, nrow = 2, scales = "free_y") + 
+  labs(x = "Year", y = "Proportion", title = "Proportion income deprived within 8km of Centre")
+
+tmp %>% filter(type == "Not Income Deprived") %>% 
+  ggplot(., 
+         aes(
+            x = year, y = prop 
+            )
+         ) + 
+  geom_line() + geom_point() +
+  facet_wrap( ~ City, nrow = 2, scales = "free_y") + 
+  labs(x = "Year", y = "Proportion", title = "Proportion Not income deprived within 8km of Centre")
+
+```
+
+# Introduction: the suburbanization of poverty
+
+Poverty has long been seen as an urban issue and in particular an ‘inner city’ problem.
+Historically, that is where older housing in worse condition tended to be found, and hence where poorer households could afford to live.
+Inner city areas were also worst hit by the processes of deindustrialization since they tended to have the oldest, least productive firms.
+There was little room for new industrial development so businesses moved out.
+More affluent households who could afford cars also moved to the suburbs and surrounding areas so the inner cities declined further.
+At the extreme, this could lead to the development of what some US commentators call ‘doughnut cities’, where a ring of relatively affluent suburbs surround an increasingly abandoned, hollowed-out core.
+
+In many European countries, the contrast between inner and outer urban areas was rarely as great, thanks in part to the efforts of planners to limit the outward sprawl of the cities but also to the scale of social policies pursued by European states.
+Spending on welfare benefits and on public services ensured that inner urban areas retained relatively strong and well-functioning communities.
+
+In recent decades, inner city locations in many developed countries have witnessed a revival their fortunes and hence in demand for housing from more affluent groups.
+This is partly about economic changes.
+With manufacturing replaced by services as the basis of the city economy, inner urban locations have not only seen an improvement in their environment, but also now offer better access to higher-skilled employment as well as expanding leisure opportunities.
+Inner city living has become the basis of a desirable lifestyle.
+
+Governments have often played a strong supporting role in these processes, investing public resources through urban regeneration schemes and seeking to leverage additional private sector investment.
+They have sought to limit suburbanization and direct new housing development towards brownfield land.
+In many cases, these policies were driven by a vision of the “sustainable city”, based around higher densities, effective public transport and
+ hoped these changes would bring new opportunities for more deprived communities as well as improvements in the physical environment.
+They sought to grow the inner city population by developing vacant or underused land, producing mixed communities.
+
+However, critics argue that these programmes have resulted not in urban regeneration where benefits are shared but in ‘gentrification’ and the displacement of poorer communities.
+The rising value of housing and the associated changes in local facilities may push out the lower income households who have long lived in these areas.
+Working class communities which had been a source of support and solidarity for their members are eroded.
+In short, the critics accuse governments of pursuing a form of “social cleansing” where the endpoint is an inner city where only the rich can afford to live while the poor are dispersed to suburban and ex-urban locations, with worse access to employment.
+
+## Suburbanisation of poverty in Scotland
+
+These processes have been little studied in Scotland.
+In 2014, however, Kavanagh, Lee & Pryce produced an AQMeN briefing paper Poverty in Suburbia: Has Glasgow gone the way of American cities?
+The paper had a strong methodological focus.
+To capture decentralisation, they used a measure called the Relative Centralisation Index (RCI) and they developed a new way of assessing uncertainty in this, using a Bayesian spatial statistics approach.
+This paper argued that there has been a decentralisation or suburbanization of poverty in Glasgow between 2001 and 2011, and that the changes were statistically significant, i.e.
+the differences were more than we would expect from the random fluctuations which will always occur in a measure from year to year.
+
+The briefing paper measured 'poverty' in three ways, each relating to specific UK state benefit receipts for persons of working age who were not in employment:
+
+1.	The proportion of the local population receiving Income Support (IS)
+2.	The proportion of the local population receiving Incapacity Benefit (IB)
+3.	The proportion of the local population receiving Jobseekers' Allowance (JSA)
+
+These analyses were expanded to the four major Scottish cities - Aberdeen, Dundee, Edinburgh and Glasgow - for a paper to appear in Annals of the American Association of Geographers, by the same authors.
+This paper calculated 95% credible intervals (CrIs) for the difference in RCI from 2001 to 2011 for the same three measures of poverty, and for each of the four cities.
+It found falls in RCI for all benefit types across all four cities, all of which were statistically significant.
+
+These papers were important for developing a new approach to the measurement of decentralization which captured the uncertainty inherent in the RCI measure.
+They were also valuable for drawing attention to a phenomenon which has been identified in other countries, notably the US, but which had not previously been discussed in the Scottish context.
+However, they can be criticized for using relatively weak proxies for poverty; in particular, they ignore those who are in work but in poverty, now the majority of working-age poor.
+They only cover two time points.
+And they do not shed any light on whether suburbanization is the result of urban regeneration and the growth of the non-poor population or whether it is the result of displacement.
+
+This paper goes beyond the earlier analysis in three ways.
+First, it uses a more sophisticated measure to identify those in poverty: the Income Deprivation measure within the Scottish Indices of Multiple Deprivation (SIMD).
+This does not rely on individual ‘out of work’ benefits for the working age but captures people in low-income households of all ages, in- or out-of-work, where they are in receipt of benefits or tax credits.
+Second, it uses the SIMD data to provide more temporal detail, tracing changes across five time periods from 2004 to 2016 to see if there is continuity in changes.
+Third, it explores the processes by which decentralization is occurring.
 
@@
 
-The purposes of this briefing paper are two-fold:
+The purposes of this briefing paper are:
 
-1.
-To reproduce the analysis in Kavanagh, Lee & Pryce 2016 using the proportion of the population 'income deprived' between 2004 and 2016, using Scottish Index of Multiple Deprivation (SIMD) definitions of income deprivation.
+1.	To reproduce the analysis in Kavanagh, Lee & Pryce 2016 using the proportion of the population 'income deprived' between 2004 and 2016, using Scottish Index of Multiple Deprivation (SIMD) definition.
 This will allow the dependence of Kavanagh, Lee & Pryce's conclusions on the measures and time periods used to be assessed.
-2.
-To use descriptive statistics and data visualisation to better understand the contribution of both changes in the numerators (population counts 'in poverty') and changes in denominators (total population counts) to overall falls in RCI.
+2.	To use descriptive statistics and data visualisation to better understand the contribution of changes in both the poor and the non-poor populations to overall changes in RCI.
 
@@
 
-The SIMD is used by the Scottish Government to identify places suffering from deprivation.
-Different editions of the SIMD have been produced in 2004, 2006, 2012, with a new release scheduled for August 2016, and made available at the small area known as the datazone, equivalent to the Lower Super Output Area (LSOA) in England & Wales.
-The methodology differs slightly from one edition to the next, but for all releases estimates of the number of individuals within each datazone who are 'income deprived', as well as overall and domain-specific scores and rankings for deprivation.
-The definition of 'income deprived' varies between household types and age groups, to account for the different needs of different demographic groups and the efficiency gains associated with living in multi-person rather than single person households.
-SIMD data releases for 2004, 2006, 2012 and 2016 were downloaded from the [SIMD website](http://www.gov.scot/Topics/Statistics/SIMD/DataAnalysis).
-The total population counts, and the population count of the subpopulation defined as 'income deprived', were extracted from each release and combined in a single dataset.
-The 'income deprived' and 'total' population counts were the focus of this report as the RCI measure, like many other segregation measures, involves comparing two mutually exclusive groups.
+The SIMD is used by the Scottish Government to identify places suffering from multiple deprivation: i.e.
+a concentration of low income households and associated problems of poor environment, poor housing, crime and social disadvantage.
+Different editions of the SIMD have been produced in 2004, 2006, 2009 and 2012, with the latest release on 31 August 2016.
+Data are made available for the small area units known as Datazones, equivalent to the Lower Super Output Area (LSOA) in England & Wales.
+For the period 2004 to 2012, the Datazone boundaries were those constructed in 2001.
+For 2016, the SIMD moved over to using the new (2011) Datazone boundaries.
+The methodology for constructing the SIMD has been consistent over this period although details differ slightly from one edition to the next.
+As well as producing an overall deprivation score and ranking for each Datazone, all five releases estimate the number of individuals within each Datazone who are 'Income Deprived', i.e.
+living in a household in receipt of one of a number of low-income benefits or tax credits (and in some cases, subject to a further low income test).
+Precise definitions change due to changes in the benefit system, but the concept and aim of the measure has remained quite consistent.
+One year (2006) is perhaps slightly different due to the omission of households in receipt of tax credits on that occasion so results for that year should perhaps be treated with more caution.
+In a recent assessment, Bailey et al (2016) have shown that there is no evidence of any urban-rural bias in this measure. 
+SIMD data releases for 2004, 2006, 2009, 2012 and 2016 were downloaded from the SIMD website.
+For each Datazone, the total population count and the Income Deprived count were extracted from each release and combined in a single dataset, and the number not Income Deprived was calculated.
+The 'Income Deprived' and ‘not Income Deprived' population counts are the focus of this report.
 
@@
 
-Producing the results involved doing the following:
+The analysis involved the following preparatory stages:
 
-* Defining centroids for each place.
-* Identifying the closest centroid for each datazone.
-* Reweighting income deprived and total population counts from the SIMD, released at 2001 datazones, onto 2011 datazones.
-* etc
-* etc
+*	Allocating Datazones to cities by defining a centroid for city, identifying the closest centroid for each Datazone and then applying a cut-off distance of around 18-20kms.
 
-### Calculating nearest centres
+*	Converting data for 2004 to 2012 SIMDs from 2001 Datazones to 2011 Datazones.
+
+### Allocating Datazones to cities
 
@@
 
-Kavanagh, Lee & Pryce (2016) produced a sensitivity analysis to assess the dependence of their results on the choice of centre, indicating they were robust to this choice.
-Alternative datazone centres could be used to explore the effect of definition for these analyses too.
+Kavanagh, Lee & Pryce (2016) produced a sensitivity analysis to assess the dependence of their results on the choice of city centroid. They found that results were robust to this choice.
 
-The majority of the results presented will be for the four cities Aberdeen, Dundee, Edinburgh and Glasgow, as these are the largest and most centralised of the areas considered.
-Within this Glasgow is historically the most populous, has the highest population density, and has the highest concentration of deprived areas.
-
-### Apportioning nearest centre to datazones
+The majority of the results presented will be for the four cities (Aberdeen, Dundee, Edinburgh and Glasgow), as these are the largest and most centralised of the areas considered.
+Within this Glasgow is the most populous, has the highest population density, and has the highest concentration of deprived areas.
 
@@
 
-As this map indicates places in the Orkney and Shetland Islands, and elsewhere in the remote and rural highlands and islands, are arbitrily assigned to either Aberdeen or Inverness, even though they are not part of any meaningful functional urban geography.
-However these are easily identifiable through their absolute distance from either centre, and are not part of the main analysis.
+As this map indicates, places in the Orkney and Shetland Islands, and elsewhere in the remote and rural highlands and islands, get assigned to either Aberdeen or Inverness, even though they are not part of any meaningful functional urban geography.
+However these areas are easily identifiable through their absolute distance from either centre.
+For the analyses below, we restrict our definition of cities to Datazones within around 20km of the city centroid.
+(Strictly speaking, we start with a cut-off of 18km but allow this to expand where this process results in ‘islands’ or individual Datazones disconnected from the rest of the city.)
 
-### Linking 2001 to 2011 datazones
+### Converting from 2001 to 2011 Datazones
 
-Income deprived and total population counts were reweighted from 2001 to 2011 datazone boundaries using a lookup file developed by [Paul Norman](http://www.geog.leeds.ac.uk/people/p.norman).
-This allowed weights to be produced, based on census output area overlap, allowing weighted 2011 population count estimates to be produced from counts enumerated at 2001 datazone level, based on the the level of overlap between 2001 and 2011 datazones.
-Further details about the lookup file was produced are available from [Paul Norman](p.d.norman@leeds.ac.uk) and the code used to perform the reweighting is reproduced below.
+[was this done after the previous stage or before? Would have thought before]
+
+Income deprived and total population counts were converted from 2001 to 2011 Datazone boundaries using a lookup file developed by [Paul Norman](http://www.geog.leeds.ac.uk/people/p.norman).
+That allowed weights to be produced, based on census output area overlap, allowing weighted 2011 population count estimates to be produced from counts enumerated at 2001 datazone level, based on the level of overlap between 2001 and 2011 datazones.
+Further details about the lookup file are available from [Paul Norman](p.d.norman@leeds.ac.uk) and the code used to perform the reweighting is reproduced below.
 
@@
 
-## RCI calculations
+To begin with, we produce maps to look at how the spatial distribution of poor relative to non-poor has changed over time.
+We also produce scatterplots to see how changes vary between more or less deprived Datazones.
 
-This section will present the results of the RCI calculations, i.e. more formal attempt to replicate the results of Kavanagh, Pryce & Lee (2016) using a different indicator and data source.
+To address the first research question, we then calculate two simple summary measures to capture changes in the distribution of poor and non-poor groups over time. These are: 
+*	The Relative Centralisation Index (RCI): this ranges from -1 to +1, with values greater than zero indicating that the poor group live, on average, closer to the centre of the city than the non-poor group.
+*	The Dissimilarity Index (D): this ranges from 0 to 1, with higher values indicating a greater tendency for poor and non-poor groups to live in different areas.
 
-### RCI, Aberdeen
+There are SIMD releases for 2004, 2006, 2009, 2012, and 2016, meaning in RCI and D scores within the period 2001 and 2011 can be observed.
+The Bayesian spatial modelling approach used by Kavanagh, Lee & Pryce (2016) was applied to each of the SIMD releases, and median RCI and D estimates, along with 95% Credible Intervals (CrIs), were produced for each release and for each city.
 
-```{r, eval = TRUE, echo = FALSE, warning = FALSE, results = 'asis'}
-posterior_summaries %>% 
-  ungroup() %>% 
-  filter(place == "Aberdeen", measure == "RCI") %>% 
-  gather(key = "bound", value = "value", lower:upper) %>% 
-  select(period, bound, value) %>% 
-  mutate(value = round(value, 3)) %>% 
-  spread(bound, value) %>% 
-  select(period, median, lower, upper) %>% 
-  kable(., format = "markdown", caption = "RCI, Aberdeen")
-
-```
-
-### RCI, Dundee
-
-```{r, eval = TRUE, echo = FALSE, warning = FALSE, results = 'asis'}
-posterior_summaries %>% 
-  ungroup() %>% 
-  filter(place == "Dundee", measure == "RCI") %>% 
-  gather(key = "bound", value = "value", lower:upper) %>% 
-  select(period, bound, value) %>% 
-  mutate(value = round(value, 3)) %>% 
-  spread(bound, value) %>% 
-  select(period, median, lower, upper) %>% 
-  kable(., format = "markdown", caption = "RCI, Dundee")
-
-```
-
-### RCI, Edinburgh
-
-```{r, eval = TRUE, echo = FALSE, warning = FALSE, results = 'asis'}
-posterior_summaries %>% 
-  ungroup() %>% 
-  filter(place == "Edinburgh", measure == "RCI") %>% 
-  gather(key = "bound", value = "value", lower:upper) %>% 
-  select(period, bound, value) %>% 
-  mutate(value = round(value, 3)) %>% 
-  spread(bound, value) %>% 
-  select(period, median, lower, upper) %>% 
-  kable(., format = "markdown", caption = "RCI, Edinburgh")
-
-```
-
-### RCI, Glasgow
-
-```{r, eval = TRUE, echo = FALSE, warning = FALSE, results = 'asis'}
-posterior_summaries %>% 
-  ungroup() %>% 
-  filter(place == "Glasgow", measure == "RCI") %>% 
-  gather(key = "bound", value = "value", lower:upper) %>% 
-  select(period, bound, value) %>% 
-  mutate(value = round(value, 3)) %>% 
-  spread(bound, value) %>% 
-  select(period, median, lower, upper) %>% 
-  kable(., format = "markdown", caption = "RCI, Glasgow")
-
-```
-
-## Dissimilarity calculations
-
-This section will present the results of the Dissimilarity index calculations, using the Bayesian modelling approach used in Kavanagh et al 2016.
-
-### D, Aberdeen
-
-```{r, eval = TRUE, echo = FALSE, warning = FALSE, results = 'asis'}
-posterior_summaries %>% 
-  ungroup() %>% 
-  filter(place == "Aberdeen", measure == "D") %>% 
-  gather(key = "bound", value = "value", lower:upper) %>% 
-  select(period, bound, value) %>% 
-  mutate(value = round(value, 3)) %>% 
-  spread(bound, value) %>% 
-  select(period, median, lower, upper) %>% 
-  kable(., format = "markdown", caption = "D, Aberdeen")
-
-```
-
-### D, Dundee
-
-```{r, eval = TRUE, echo = FALSE, warning = FALSE, results = 'asis'}
-posterior_summaries %>% 
-  ungroup() %>% 
-  filter(place == "Dundee", measure == "D") %>% 
-  gather(key = "bound", value = "value", lower:upper) %>% 
-  select(period, bound, value) %>% 
-  mutate(value = round(value, 3)) %>% 
-  spread(bound, value) %>% 
-  select(period, median, lower, upper) %>% 
-  kable(., format = "markdown", caption = "D, Dundee")
-
-```
-
-### D, Edinburgh
-
-```{r, eval = TRUE, echo = FALSE, warning = FALSE, results = 'asis'}
-posterior_summaries %>% 
-  ungroup() %>% 
-  filter(place == "Edinburgh", measure == "D") %>% 
-  gather(key = "bound", value = "value", lower:upper) %>% 
-  select(period, bound, value) %>% 
-  mutate(value = round(value, 3)) %>% 
-  spread(bound, value) %>% 
-  select(period, median, lower, upper) %>% 
-  kable(., format = "markdown", caption = "D, Edinburgh")
-
-```
-
-### RCI, Glasgow
-
-```{r, eval = TRUE, echo = FALSE, warning = FALSE, results = 'asis'}
-posterior_summaries %>% 
-  ungroup() %>% 
-  filter(place == "Glasgow", measure == "D") %>% 
-  gather(key = "bound", value = "value", lower:upper) %>% 
-  select(period, bound, value) %>% 
-  mutate(value = round(value, 3)) %>% 
-  spread(bound, value) %>% 
-  select(period, median, lower, upper) %>% 
-  kable(., format = "markdown", caption = "D, Glasgow")
-
-```
-
-### Change in RCI and D over time
-
-Unlike the census-based variables used in Kavanagh, Lee & Pryce (2016), there are SIMD releases for 2004, 2006, 2009, 2012, and 2016, meaning in RCI and D scores within the period 2001 and 2011 can be observed.
-The Bayesian spatial modelling approach used by Kavanagh, Lee & Pryce (2016) was applied to each of the SIMD releases, and median RCI and D estimates, along with 95% CrIs, were produced for each release and for each city.
-
-The graph of RCI over time is shown below.
-The density of RCI estimates for each place and in each year is shown using a violin plot, with longer shapes indicating a greater range of posterior estimates.
-The median RCI scores are indicated as points, linked for each city with dashed lines.
-
-```{r plot_change_rci, eval = TRUE, echo = FALSE, warning = FALSE}
-# Do RCI graph
-
-all_posteriors  %>% 
-  filter(measure == "RCI")   %>% 
-  group_by(place, period)  %>% 
-  mutate(med_val = median(value))  %>%  
-  mutate(pp = paste0(place, period))  %>% 
-  ggplot(.) + 
-  geom_violin(
-    aes(x = factor(period), y = value, group = pp, fill = place), 
-    alpha = 0.2, position = "identity"
-  ) + 
-  geom_line(
-    aes(x = factor(period), y = med_val, group = place, colour = place), 
-    size = 1.1, linetype = "dashed"
-  ) + 
-  geom_point(
-    aes(x = factor(period), y = med_val, shape = place, colour = place), 
-    size = 3.5
-  ) + 
-  labs(
-    x = "Year", y = "RCI", colour = "City", 
-    shape = "City", fill = "City", title = "Change in RCI scores over time"
-  ) + 
-  geom_hline(aes(yintercept = 0))
-
-```
-
-The overall trend in each place is for RCI to reduce from 2004 to 2012.
-However within each period the ordering of RCIs between places has not changed, with Dundee showing the most centralisation of poverty, then Aberdeen, and Glasgow, with Edinburgh the most decentralised.
-In terms of trends over time, Glasgow and Dundee both show steadily declining RCI between SIMD releases, whereas for Aberdeen and Edinbugh RCI values increased from 2004 to 2006, before declining in 2009 and 2012.
-From 2009 onwards Edinburgh has negative RCI scores, suggesting a peripheralisation rather than centralisation of poverty.
-
-The equivalent graph for the Dissimilarity index is shown below:
-
-```{r plot_change_d, eval = TRUE, echo = FALSE, warning = FALSE}
-# Do D graph
-
-all_posteriors %>% 
-  filter(measure == "D")   %>% 
-  group_by(place, period)  %>% 
-  mutate(med_val = median(value))  %>%  
-  mutate(pp = paste0(place, period))  %>% 
-  ggplot(.) + 
-  geom_violin(
-    aes(x = factor(period), y = value, group = pp, fill = place), 
-    alpha = 0.2, position = "identity"
-  ) + 
-  geom_line(
-    aes(x = factor(period), y = med_val, group = place, colour = place), 
-    size = 1.1, linetype = "dashed"
-  ) + 
-  geom_point(
-    aes(x = factor(period), y = med_val, shape = place, colour = place), 
-    size = 3.5
-  ) + 
-  labs(
-    x = "Year", y = "D", colour = "City", shape = "City", 
-    fill = "City", title = "Change in D scores over time")
-
-```
-
-As with RCI, Dissimilarity scores have also fallen over time for all four cities.
-The greatest contrast in the trends is between Aberdeen and Edinburgh.
-In 2004 both cities had similar and overlapping D scores, but between 2004 and 2009 D fell rapidly for Edinburgh, while remaining fairly stable in Aberdeen, before falling from 2009 to 2012.
-Glasgow has always had the lowest D scores of the four cities, and Dundee the highest, but since 2009 the distribution of D estimates for Edinburgh have started to overlap those of Glasgow as the rate of decline in D has been slightly greater in Edinburgh than Glasgow.
-Similarly, the the greatest fall in D occurred between 2006 and 2009.
+To address the second research question, we look at changes in the number of people in the poor and non-poor groups by distance from the city centre.
 
@@
 
-The following map shows the proportion of the population who are classified as income deprived for each datazone, in the years 2004, 2006, 2009 and 2012.
+The maps shows the proportion of the population who are classified as Income Deprived for each Datazone, in the years 2004 to 2016.
 Given that so much of Scotland is rural, few areas appear income deprived using a standard choropleth.
@@
 The map below focuses in on income deprivation within Glasgow, and in particular within datazones not more than 15 km from Glasgow city centre.
+From this map there are suggestions of a decentralisation of poverty, in particular between 2009 and 2012.
+However there are also clear pockets of areas, in the south, east and north west of Glasgow city, with persistently high levels of income deprivation.
 
@@
 
-An equivalent map for Edinburgh is produced below:
+An equivalent map for Edinburgh is produced below.
+This map also suggests that, as in Glasgow, drivers of increasing levels of decentralisation of poverty in Edinburgh include both rapid falls in poverty in the centre, alongside some more suburban areas with persistently high levels of poverty.
+In Edinburgh's case, in pockets towards the West, North West, and South/South East of the city centre.
 
@@
 
-This map also suggests that, as in Glasgow, drivers of increasing levels of decentralisation of poverty in Edinburgh include both rapid falls in poverty in the centre, alongside some more suburban areas with persistently high levels of poverty.
-In Edinburgh's case, in pockets towards the West, North West, and South/South East of the city centre.
-
-The equivalent map for Dundee is shown below:
+The equivalent map for Dundee is shown below.
+Within Dundee the patterns of spatial persistence of poverty appear most consistent over time, with a clearer demarcation between urban and non-urban areas.
+The functional urban geography of Dundee appears more spatially concentrated than for Glasgow and Edinburgh.
 
@@
 
-Within Dundee the patterns of spatial persistence of poverty appear most consistent over time, with a clearer demarcation between urban and non-urban areas.
-The functional urban geography of Dundee appears more spatially concentrated than for Glasgow and Edinburgh.
-
-The equivalent map for Aberdeen, historically the most affluent and rural of the four cities, is shown here:
+The equivalent map for Aberdeen, historically the most affluent and rural of the four cities, is shown here.
+As with Glasgow and Edinburgh, it appears both that the greatest changes have occurred between 2009 and 2012, and also that any fall in RCI scores are likely due to persistence of poverty in some local areas, with a high income deprivation area clearly visible in all years to the north west of the city centre.
 
@@
 ```
-
-As with Glasgow and Edinburgh, it appears both that the greatest changes have occurred between 2009 and 2012, and also that any fall in RCI scores are likely due to persistence of poverty in some local areas, with a high income deprivation area clearly visible in all years to the north west of the city centre.
 
@@
 
-## Table of average proportion income deprived by year
+## Income Deprivation by city
 
@@
 
+## RCI calculations
+
+This section presents the results of the RCI calculations, using the Bayesian modelling approach used in Kavanagh et al 2016.
+
+### RCI, Aberdeen
+
+```{r, eval = TRUE, echo = FALSE, warning = FALSE, results = 'asis'}
+posterior_summaries %>% 
+  ungroup() %>% 
+  filter(place == "Aberdeen", measure == "RCI") %>% 
+  gather(key = "bound", value = "value", lower:upper) %>% 
+  select(period, bound, value) %>% 
+  mutate(value = round(value, 3)) %>% 
+  spread(bound, value) %>% 
+  select(period, median, lower, upper) %>% 
+  kable(., format = "markdown", caption = "RCI, Aberdeen")
+
+```
+
+### RCI, Dundee
+
+```{r, eval = TRUE, echo = FALSE, warning = FALSE, results = 'asis'}
+posterior_summaries %>% 
+  ungroup() %>% 
+  filter(place == "Dundee", measure == "RCI") %>% 
+  gather(key = "bound", value = "value", lower:upper) %>% 
+  select(period, bound, value) %>% 
+  mutate(value = round(value, 3)) %>% 
+  spread(bound, value) %>% 
+  select(period, median, lower, upper) %>% 
+  kable(., format = "markdown", caption = "RCI, Dundee")
+
+```
+
+### RCI, Edinburgh
+
+```{r, eval = TRUE, echo = FALSE, warning = FALSE, results = 'asis'}
+posterior_summaries %>% 
+  ungroup() %>% 
+  filter(place == "Edinburgh", measure == "RCI") %>% 
+  gather(key = "bound", value = "value", lower:upper) %>% 
+  select(period, bound, value) %>% 
+  mutate(value = round(value, 3)) %>% 
+  spread(bound, value) %>% 
+  select(period, median, lower, upper) %>% 
+  kable(., format = "markdown", caption = "RCI, Edinburgh")
+
+```
+
+### RCI, Glasgow
+
+```{r, eval = TRUE, echo = FALSE, warning = FALSE, results = 'asis'}
+posterior_summaries %>% 
+  ungroup() %>% 
+  filter(place == "Glasgow", measure == "RCI") %>% 
+  gather(key = "bound", value = "value", lower:upper) %>% 
+  select(period, bound, value) %>% 
+  mutate(value = round(value, 3)) %>% 
+  spread(bound, value) %>% 
+  select(period, median, lower, upper) %>% 
+  kable(., format = "markdown", caption = "RCI, Glasgow")
+
+```
+
+## Dissimilarity calculations
+
+This section presents the results of the Dissimilarity index calculations, using the Bayesian modelling approach used in Kavanagh et al 2016.
+
+### D, Aberdeen
+
+```{r, eval = TRUE, echo = FALSE, warning = FALSE, results = 'asis'}
+posterior_summaries %>% 
+  ungroup() %>% 
+  filter(place == "Aberdeen", measure == "D") %>% 
+  gather(key = "bound", value = "value", lower:upper) %>% 
+  select(period, bound, value) %>% 
+  mutate(value = round(value, 3)) %>% 
+  spread(bound, value) %>% 
+  select(period, median, lower, upper) %>% 
+  kable(., format = "markdown", caption = "D, Aberdeen")
+
+```
+
+### D, Dundee
+
+```{r, eval = TRUE, echo = FALSE, warning = FALSE, results = 'asis'}
+posterior_summaries %>% 
+  ungroup() %>% 
+  filter(place == "Dundee", measure == "D") %>% 
+  gather(key = "bound", value = "value", lower:upper) %>% 
+  select(period, bound, value) %>% 
+  mutate(value = round(value, 3)) %>% 
+  spread(bound, value) %>% 
+  select(period, median, lower, upper) %>% 
+  kable(., format = "markdown", caption = "D, Dundee")
+
+```
+
+### D, Edinburgh
+
+```{r, eval = TRUE, echo = FALSE, warning = FALSE, results = 'asis'}
+posterior_summaries %>% 
+  ungroup() %>% 
+  filter(place == "Edinburgh", measure == "D") %>% 
+  gather(key = "bound", value = "value", lower:upper) %>% 
+  select(period, bound, value) %>% 
+  mutate(value = round(value, 3)) %>% 
+  spread(bound, value) %>% 
+  select(period, median, lower, upper) %>% 
+  kable(., format = "markdown", caption = "D, Edinburgh")
+
+```
+
+### D, Glasgow
+
+```{r, eval = TRUE, echo = FALSE, warning = FALSE, results = 'asis'}
+posterior_summaries %>% 
+  ungroup() %>% 
+  filter(place == "Glasgow", measure == "D") %>% 
+  gather(key = "bound", value = "value", lower:upper) %>% 
+  select(period, bound, value) %>% 
+  mutate(value = round(value, 3)) %>% 
+  spread(bound, value) %>% 
+  select(period, median, lower, upper) %>% 
+  kable(., format = "markdown", caption = "D, Glasgow")
+
+```
+
+### Change in RCI and D over time
+
+The graph of RCI over time is shown below.
+The density of RCI estimates for each place and in each year is shown using a violin plot, with longer shapes indicating a greater range of posterior estimates.
+The median RCI scores are indicated as points, linked for each city with dashed lines.
+
+```{r plot_change_rci, eval = TRUE, echo = FALSE, warning = FALSE}
+# Do RCI graph
+
+all_posteriors  %>% 
+  filter(measure == "RCI")   %>% 
+  group_by(place, period)  %>% 
+  mutate(med_val = median(value))  %>%  
+  mutate(pp = paste0(place, period))  %>% 
+  ggplot(.) + 
+  geom_violin(
+    aes(x = factor(period), y = value, group = pp, fill = place), 
+    alpha = 0.2, position = "identity"
+  ) + 
+  geom_line(
+    aes(x = factor(period), y = med_val, group = place, colour = place), 
+    size = 1.1, linetype = "dashed"
+  ) + 
+  geom_point(
+    aes(x = factor(period), y = med_val, shape = place, colour = place), 
+    size = 3.5
+  ) + 
+  labs(
+    x = "Year", y = "RCI", colour = "City", 
+    shape = "City", fill = "City", title = "Change in RCI scores over time"
+  ) + 
+  geom_hline(aes(yintercept = 0))
+
+```
+
+[CHECK] The overall trend for all four cities is for poverty to decentralize between 2004 to 2016.
+Dundee shows the most centralisation of poverty, then Aberdeen, and Glasgow, with Edinburgh the most decentralised.
+Indeed, from 2009 onwards, Edinburgh is unique in having poorer households living further from the city centre than non-poor.
+
+[CHECK] For Glasgow and Dundee, the decline in RCI is continuous, whereas for Aberdeen and Edinbugh RCI values increased from 2004 to 2006, before declining in 2009 and 2012.
+As note above, there is some reason to be more cautious about the comparability of the 2006 data with that from other years.
+
+The equivalent graph for the Dissimilarity index is shown below:
+
+```{r plot_change_d, eval = TRUE, echo = FALSE, warning = FALSE}
+# Do D graph
+
+all_posteriors %>% 
+  filter(measure == "D")   %>% 
+  group_by(place, period)  %>% 
+  mutate(med_val = median(value))  %>%  
+  mutate(pp = paste0(place, period))  %>% 
+  ggplot(.) + 
+  geom_violin(
+    aes(x = factor(period), y = value, group = pp, fill = place), 
+    alpha = 0.2, position = "identity"
+  ) + 
+  geom_line(
+    aes(x = factor(period), y = med_val, group = place, colour = place), 
+    size = 1.1, linetype = "dashed"
+  ) + 
+  geom_point(
+    aes(x = factor(period), y = med_val, shape = place, colour = place), 
+    size = 3.5
+  ) + 
+  labs(
+    x = "Year", y = "D", colour = "City", shape = "City", 
+    fill = "City", title = "Change in D scores over time")
+
+```
+
+As with RCI, Dissimilarity scores have also fallen over time for all four cities.
+The greatest contrast in the trends is between Aberdeen and Edinburgh.
+In 2004 both cities had similar and overlapping D scores, but between 2004 and 2009 segregation fell rapidly for Edinburgh, while remaining fairly stable in Aberdeen, before falling from 2009 to 2012.
+Glasgow has always had the lowest D scores of the four cities, and Dundee the highest, but since 2009 the distribution of D estimates for Edinburgh have started to overlap those of Glasgow as the rate of decline in D has been slightly greater in Edinburgh than Glasgow.
+Similarly, the the greatest fall in D occurred between 2006 and 2009.
+
+In general, this suggests segregation is falling or social mix increasing.
+However, given the findings about about the decentralization of poverty, the concern is that this reduction is merely a temporary phenomenon during a period of structural change: as non-poor groups move back to colonise inner city locations, mixing increases but is then likely to fall again as the poor continue to be displaced.
+
 ## Plots of the distance/deprivation relationship
 
-In the figure below, the vertical axis shows the proportion change in income deprivation from 2004 to 2012, and the horizontal axis shows the distance of datazones from the city centres, plotted on a log scale, for each of the four main Scottish cities.
+[CHECK] In the figure below, the vertical axis shows the change in the proportion income deprived from 2004 to 2016, and the horizontal axis shows the distance of datazones from the city centres, plotted on a log scale, for each of the four main Scottish cities.
 Individual datazones are indicated with faint circles, and a nonlinear smoother has been added showing how this the change in income deprivation varies as a function of distance.
@@
 These plots suggest a qualitatively similar relationship between change in deprivation and distance, up to around 12km from city centre, for Aberdeen, Edinburgh and Glasgow, with Glasgow showing the strongest gradient of change between these two periods.
-Dundee shows a simpilar kind of relationship, but concentrated within the first 5km, rather than 12km, of distance from city centre.
+Dundee shows a similar kind of relationship, but concentrated within the first 5km, rather than 12km, of distance from city centre.
 This is consistent with the maps of income deprivation in Dundee, showing much greater levels of spatial concentration than for the other cities.
@@
 The falls in income deprivation nearer city centres are a function both of changing numerators - i.e.
-changes to income deprived population sizes - and changing denominators - i.e.
+changes to the number of people income deprived - and changing denominators - i.e.
 changes to total population sizes.
@@
 ```
-
 This plot indicates proportionate falls in the ID population shares at distances closes to the city centre, up to around 20% in Aberdeen and Glasgow, and either proportionate rises (Dundee and Edinburgh) or little change in proportions (Aberdeen and Glasgow) at greater distance from the city centre.
@@
 
-The following plot attempts to summarise the previous series of results, in order the represent how both total population sizes and income deprived population sizes have changed with distance, and the overall strength of the relationship between income deprived population share and distance from city centre.
-
-For each of the four cities, separate linear regressions were produced to extract the gradient of changes in total population size, income deprived population size, and changes in income deprived population shares with distance, over distances between 1km and 12km.
-The horizontal axis shows the gradient of change in total population with distance, the vertical axis shows graident of change in income deprived population with distance.
-Each city is represented by a separate point, labelled G for Glasgow, A for Aberdeen, D for Dundee and E for Edinburgh.
-The darkness of these symbols indicates the magnitude of the overall change in the income deprived population shares as a function of distance, with darker shades indicating a larger relationship and lighter shades a smaller relationship.
-
-```{r plot_all_bubble, eval = TRUE, echo = FALSE , message = FALSE, warning = FALSE}
-
-source("scripts/produce_dz2011_regression_bubble.R")
-
-# As a scatterplot?
-
-simple_results %>% 
-  mutate(place2 = str_sub(place, 1, 1)) %>% 
-  select(type, place2, estimate) %>% 
-  spread(type, estimate) %>% 
-  ggplot(., aes(x = denominator, y = numerator)) + 
-  geom_point(size = 10, colour = "lightblue") +
-  geom_text(aes(label = place2, alpha =  propid), show.legend = F) +
-  coord_cartesian(xlim = c(-0.1, 0.1), ylim = c(0.0, 0.1)) +
-  geom_hline(yintercept = 0, linetype = "dashed") + 
-  geom_vline(xintercept = 0, linetype = "dashed") + 
-  labs(x = "change in total population with distance", y = "change in income deprived population with distance")
-
-```
-
-# Discussion
-
-## Substantive Findings
-
-The apparent fall in income deprivation proportions in more central areas of cities, in particular Glasgow, and in particular the fact the fall appeared greatest between 2008 and 2012, suggests there may be issues with the way in which income deprivation has been defined, and how effectively or otherwise this connects to broader and alternative definitions of poverty other this period.
-Although the definition of income deprived is not completely consistent from one SIMD release to the next, they tend to be based on claims for particular types of state benefit.
-During the 2008 to 2012 period the level of conditionality associated with many benefits increased for people of working age, while at the same time the longest sustained fall and plateau in real wages in over half a century occurred.
-Unemployment rates did not rise as they had in previous recessions, but levels of self-employment, short-term contracts, and part time work all increased.
-The self-employed, in particular, are not eligible for many of the benefits that people earning low wages as employees receive, and so the fall the income deprivation may reflect this compositional change in the workforce.
-Similarly increasing casualisation of employment, with variable hours and pay from one week to the next, may put off many people earning low wages on average from applying for 'top up' benefits such as tax credits.
-Again, indications that there have been falls in Scotland in income deprivation between 2008 to 2012 suggest statistical artefacts.
-
-## Limitations and recommendations for further research
+# Limitations and recommendations for further research
 

### 2016-08-24 scripts/from_gavin/do_rci_and_d_for_all_permutations.R

@@ -77,14 +77,16 @@ do_model <- function(place, initial_dist = 18000){
     as.integer(dz_city@data$pop_id_2004),
     as.integer(dz_city@data$pop_id_2006), 
     as.integer(dz_city@data$pop_id_2009), 
-    as.integer(dz_city@data$pop_id_2012) 
+    as.integer(dz_city@data$pop_id_2012),
+    as.integer(dz_city@data$pop_id_2012) # placeholder for new data
   )
   
   N.mat <- cbind(
     as.integer(dz_city@data$pop_total_2004),
     as.integer(dz_city@data$pop_total_2006), 
     as.integer(dz_city@data$pop_total_2009), 
-    as.integer(dz_city@data$pop_total_2012) 
+    as.integer(dz_city@data$pop_total_2012),
+    as.integer(dz_city@data$pop_total_2012) # placeholder for new data
   )
 
 
@@ -102,23 +104,28 @@ do_model <- function(place, initial_dist = 18000){
   Y <- Y.mat[,4] ; N <- N.mat[,4]
   model_04 <- S.CARleroux(Y ~ 1, trials = N, W = W, family = "binomial", burnin = burnin, n.sample = n.sample, thin = thin )  
   
-
+  #placeholder for new data 
+  Y <- Y.mat[,5] ; N <- N.mat[,5]
+  model_05 <- S.CARleroux(Y ~ 1, trials = N, W = W, family = "binomial", burnin = burnin, n.sample = n.sample, thin = thin )  
+  
 #  model$summary.results
   
   #### Compute the coordinates and ordering from the city centre
   dist.order <- order(dz_city@data$distance_to_centre)
   
   #### Compute the global RCI and D
-  indicators.post <- array(NA, c(n.keep,8))
+  indicators.post <- array(NA, c(n.keep,10))
   colnames(indicators.post) <- c(
-    "RCI_2004", "RCI_2006","RCI_2009","RCI_2012", 
-    "D_2004", "D_2006", "D_2009", "D_2012"
+    "RCI_2004", "RCI_2006","RCI_2009","RCI_2012", "RCI_2015", 
+    "D_2004", "D_2006", "D_2009", "D_2012", "D_2015"
   )
   
   all_2004 <- as.integer(N.mat[,1])
   all_2006 <- as.integer(N.mat[,2])
[…]

### 2016-08-31 markdown/draft.RmD
add simd 2016; redo simd code; redo model run code; redo markdown

@@
 ---
-title: "The Suburbanisation of Poverty in Scotland’s Cities: Urban Regeneration or Social Cleansing?"
+title: 'The Suburbanisation of Poverty in Scotland’s Cities: Urban Regeneration or
+  Social Cleansing?'
 author: "Jon Minton & Nick Bailey"
 date: "31 August 2016"
-output: html_document
+output: word_document
 ---
@@
 ```{r id_props, eval = TRUE, echo = FALSE, cache = FALSE, message = FALSE, warning = FALSE}
+
 simd_2011_reweighted  %>% 
@@
     filter(City %in% c("Aberdeen", "Dundee", "Edinburgh", "Glasgow")) %>% 
-    mutate(centralness = cut(
-      distance_to_centre, 
-      breaks = c(0, 8000, 20000, Inf), 
-      labels = c("centre", "periphery", "exurb"))
-    ) %>% 
-  filter(centralness != "exurb") %>% 
+    mutate(
+      lower_thresh = ifelse(City %in% c("Aberdeen", "Dundee"), 3200, 5000),
+      centralness = ifelse(distance_to_centre < 20000, "periphery", "exurb"),
+      centralness = ifelse(distance_to_centre < lower_thresh, "centre", centralness)
+      ) %>% dplyr::select(-lower_thresh) %>% 
+  filter(centralness %in% c("centre", "periphery")) %>% 
   group_by(City, year, centralness) %>%
@@
   facet_wrap( ~ City, nrow = 2, scales = "free_y") + 
-  labs(x = "Year", y = "Proportion", title = "Proportion income deprived within 8km of Centre")
+  labs(x = "Year", y = "Proportion", title = "Proportion income deprived near City Centre")
 
@@
   facet_wrap( ~ City, nrow = 2, scales = "free_y") + 
-  labs(x = "Year", y = "Proportion", title = "Proportion Not income deprived within 8km of Centre")
+  labs(x = "Year", y = "Proportion", title = "Proportion Not income deprived near City Centre")
 
@@
 
+*	Converting data for 2004 to 2012 SIMDs from 2001 Datazones to 2011 Datazones.
 *	Allocating Datazones to cities by defining a centroid for city, identifying the closest centroid for each Datazone and then applying a cut-off distance of around 18-20kms.
 
-*	Converting data for 2004 to 2012 SIMDs from 2001 Datazones to 2011 Datazones.
+### Converting from 2001 to 2011 Datazones
+
+Income deprived and total population counts were converted from 2001 to 2011 Datazone boundaries using a lookup file developed by [Paul Norman](http://www.geog.leeds.ac.uk/people/p.norman).
+That allowed weights to be produced, based on census output area overlap, allowing weighted 2011 population count estimates to be produced from counts enumerated at 2001 datazone level, based on the level of overlap between 2001 and 2011 datazones.
+Further details about the lookup file are available from [Paul Norman](p.d.norman@leeds.ac.uk) and the code used to perform the reweighting is reproduced below.
 
@@
 
-### Converting from 2001 to 2011 Datazones
-
-[was this done after the previous stage or before? Would have thought before]
-
-Income deprived and total population counts were converted from 2001 to 2011 Datazone boundaries using a lookup file developed by [Paul Norman](http://www.geog.leeds.ac.uk/people/p.norman).
-That allowed weights to be produced, based on census output area overlap, allowing weighted 2011 population count estimates to be produced from counts enumerated at 2001 datazone level, based on the level of overlap between 2001 and 2011 datazones.
-Further details about the lookup file are available from [Paul Norman](p.d.norman@leeds.ac.uk) and the code used to perform the reweighting is reproduced below.
-
 # Results
@@
 simd_2011_reweighted %>% 
-  select(dz_2011, year, pop_incomedeprived, pop_total) %>%
+  dplyr::select(dz_2011, year, pop_incomedeprived, pop_total) %>%
   mutate(prop_id = pop_incomedeprived / pop_total) %>% 
   mutate(prop_id = ifelse(prop_id > 0.5, 0.5, prop_id)) %>% 
-  select(dz_2011, year, prop_id) %>% 
+  dplyr::select(dz_2011, year, prop_id) %>% 
   spread(year, prop_id) %>% 
@@
   tm_shape(.) + 
-  tm_facets(ncol = 2) + 
+  tm_facets(ncol = 3) + 
   tm_fill(
-    col = c("2004", "2006", "2009", "2012"), 
+    col = c("2004", "2006", "2009", "2012", "2016"), 
     palette = brewer.pal(10, "Spectral"), 
@@
 simd_2011_reweighted %>% 
-  select(dz_2011, year, pop_incomedeprived, pop_total) %>%
+  dplyr::select(dz_2011, year, pop_incomedeprived, pop_total) %>%
   mutate(prop_id = pop_incomedeprived / pop_total) %>% 
   mutate(prop_id = ifelse(prop_id > 0.5, 0.5, prop_id)) %>% 
-  select(dz_2011, year, prop_id) %>% 
+  dplyr::select(dz_2011, year, prop_id) %>% 
   left_join(centre_distance, by = c("dz_2011" = "datazone")) %>% 
@@
   filter(distance_to_centre < 15000) %>% 
-  select(dz_2011, year, prop_id) %>% 
+  dplyr::select(dz_2011, year, prop_id) %>% 
   spread(year, prop_id) %>% 
@@
 ```{r show_temporal_dep, echo = FALSE, warning = FALSE, EVAL = TRUE}
-
 simd_2011_reweighted %>% 
-  filter(year %in% c(2004, 2012)) %>% 
+  filter(year %in% c(2004, 2016)) %>% 
   mutate(prop_id = pop_incomedeprived / pop_total) %>% 
-  select(dz_2011, year, prop_id) %>% 
+  dplyr::select(dz_2011, year, prop_id) %>% 
   spread(year, prop_id) %>% 
-  rename(prop_id_2004 = `2004`, prop_id_2012 = `2012`) %>% 
-  ggplot(., aes(x = prop_id_2004, y = prop_id_2012)) + 
+  rename(prop_id_2004 = `2004`, prop_id_2016 = `2016`) %>% 
+  ggplot(., aes(x = prop_id_2004, y = prop_id_2016)) + 
   geom_point(alpha = 0.1) + 
@@
   labs(
-    x = "ID prop, 2004", y = "ID prop, 2012",
-    title = "2012 against 2004"
-       ) -> s_04_12
+    x = "ID prop, 2004", y = "ID prop, 2016",
+    title = "2016 against 2004"
+       ) -> s_04_16
 
@@
   mutate(prop_id = pop_incomedeprived / pop_total) %>% 
-  select(dz_2011, year, prop_id) %>% 
+  dplyr::select(dz_2011, year, prop_id) %>% 
   spread(year, prop_id) %>% 
@@
 
-plot_grid(s_04_12, s_04_06, s_06_09, s_09_12, nrow = 2)
+simd_2011_reweighted %>% 
+  filter(year %in% c(2012, 2016)) %>% 
+  mutate(prop_id = pop_incomedeprived / pop_total) %>% 
+  dplyr::select(dz_2011, year, prop_id) %>% 
+  spread(year, prop_id) %>% 
+  rename(prop_id_2012 = `2012`, prop_id_2016 = `2016`) %>% 
+  ggplot(., aes(x = prop_id_2012, y = prop_id_2016)) + 
+  geom_point(alpha = 0.1) + 
+  geom_abline(colour = "blue", slope = 1, intercept = 0) + 
+  stat_smooth(colour = "red", linetype = "dashed") + 
+  labs(
+    x = "ID prop, 2012", y = "ID prop, 2016",
+    title = "2016 against 2012"
+  ) -> s_12_16
+
+plot_grid(s_04_16, s_04_06, s_06_09, s_09_12, s_12_16, nrow = 2)
 
@@
     mutate(prop_id = 100 * round(prop_id, 3)) %>% 
-    select(City = nearest_centre, prop_id, Year = year) %>% 
+    dplyr::select(City = nearest_centre, prop_id, Year = year) %>% 
     spread(Year, prop_id) %>% 
@@
     kable(., format = "markdown", caption = "Income deprived % by City and Year")
+
+```
+
+The size of the income deprived population is shown below:
+
+```{r id_size_by_year, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+  simd_2011_reweighted  %>% 
+    ungroup() %>%  
+    left_join(centre_distance, by = c("dz_2011" = "datazone")) %>%
+    filter(distance_to_centre < 15000) %>% 
+    group_by(nearest_centre, year) %>% 
+    summarise(pop_incomedeprived = sum(pop_incomedeprived)) %>% 
+    dplyr::select(City = nearest_centre, pop_incomedeprived, Year = year) %>% 
+   mutate(pop_incomedeprived = round(pop_incomedeprived, 0)) %>% 
+    spread(Year, pop_incomedeprived) %>% 
+    ungroup() %>% 
+    arrange(`2004`) %>% 
+    kable(., format = "markdown", caption = "Income deprived size by City and Year")
+
+```
+
+And the size of the non-income deprived population is shown below:
+
+```{r non_id_size_by_year, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+  simd_2011_reweighted  %>% 
+    ungroup() %>%  
+    left_join(centre_distance, by = c("dz_2011" = "datazone")) %>%
+    filter(distance_to_centre < 15000) %>% 
+    group_by(nearest_centre, year) %>% 
+    summarise(pop_incomedeprived = sum(pop_incomedeprived),
+              pop_total = sum(pop_total)
+              ) %>%
+    mutate(pop_nonincomedeprived = pop_total - pop_incomedeprived) %>% 
+    dplyr::select(City = nearest_centre, pop_nonincomedeprived, Year = year) %>% 
+   mutate(pop_nonincomedeprived = round(pop_nonincomedeprived, 0)) %>% 
+    spread(Year, pop_nonincomedeprived) %>% 
+    ungroup() %>% 
+    arrange(`2004`) %>% 
+    kable(., format = "markdown", caption = "Non-Income deprived size by City and Year")
 
@@
   gather(key = "bound", value = "value", lower:upper) %>% 
-  select(period, bound, value) %>% 
+  dplyr::select(period, bound, value) %>% 
   mutate(value = round(value, 3)) %>% 
   spread(bound, value) %>% 
-  select(period, median, lower, upper) %>% 
+  dplyr::select(period, median, lower, upper) %>% 
   kable(., format = "markdown", caption = "RCI, Aberdeen")
@@
   gather(key = "bound", value = "value", lower:upper) %>% 
-  select(period, bound, value) %>% 
+  dplyr::select(period, bound, value) %>% 
   mutate(value = round(value, 3)) %>% 
   spread(bound, value) %>% 
-  select(period, median, lower, upper) %>% 
+  dplyr::select(period, median, lower, upper) %>% 
   kable(., format = "markdown", caption = "RCI, Dundee")
@@
   gather(key = "bound", value = "value", lower:upper) %>% 
-  select(period, bound, value) %>% 
+  dplyr::select(period, bound, value) %>% 
   mutate(value = round(value, 3)) %>% 
   spread(bound, value) %>% 
-  select(period, median, lower, upper) %>% 
+  dplyr::select(period, median, lower, upper) %>% 
   kable(., format = "markdown", caption = "RCI, Edinburgh")
@@
   gather(key = "bound", value = "value", lower:upper) %>% 
-  select(period, bound, value) %>% 
+  dplyr::select(period, bound, value) %>% 
   mutate(value = round(value, 3)) %>% 
   spread(bound, value) %>% 
-  select(period, median, lower, upper) %>% 
+  dplyr::select(period, median, lower, upper) %>% 
   kable(., format = "markdown", caption = "RCI, Glasgow")
@@
   gather(key = "bound", value = "value", lower:upper) %>% 
-  select(period, bound, value) %>% 
+  dplyr::select(period, bound, value) %>% 
   mutate(value = round(value, 3)) %>% 
   spread(bound, value) %>% 
-  select(period, median, lower, upper) %>% 
+  dplyr::select(period, median, lower, upper) %>% 
   kable(., format = "markdown", caption = "D, Aberdeen")
@@
   gather(key = "bound", value = "value", lower:upper) %>% 
-  select(period, bound, value) %>% 
+  dplyr::select(period, bound, value) %>% 
   mutate(value = round(value, 3)) %>% 
   spread(bound, value) %>% 
-  select(period, median, lower, upper) %>% 
+  dplyr::select(period, median, lower, upper) %>% 
   kable(., format = "markdown", caption = "D, Dundee")
@@
   gather(key = "bound", value = "value", lower:upper) %>% 
-  select(period, bound, value) %>% 
+  dplyr::select(period, bound, value) %>% 
   mutate(value = round(value, 3)) %>% 
   spread(bound, value) %>% 
-  select(period, median, lower, upper) %>% 
+  dplyr::select(period, median, lower, upper) %>% 
   kable(., format = "markdown", caption = "D, Edinburgh")
@@
   gather(key = "bound", value = "value", lower:upper) %>% 
-  select(period, bound, value) %>% 
+  dplyr::select(period, bound, value) %>% 
   mutate(value = round(value, 3)) %>% 
   spread(bound, value) %>% 
-  select(period, median, lower, upper) %>% 
+  dplyr::select(period, median, lower, upper) %>% 
   kable(., format = "markdown", caption = "D, Glasgow")
@@
     mutate(prop_id = pop_incomedeprived / pop_total) %>% 
-    select(dz_2011, distance_to_centre, year, prop_id) %>% 
+    dplyr::select(dz_2011, distance_to_centre, year, prop_id) %>% 
     mutate(distance_to_centre = distance_to_centre / 1000) %>%
+    filter(distance_to_centre < 30) %>% 
     spread(year, prop_id) %>% 
-    mutate(change = `2012` - `2004`) %>% 
+    mutate(change = `2016` - `2004`) %>% 
     ggplot(., aes(x = distance_to_centre, y = change)) + 
     geom_point( alpha = 0.01) + 
-    scale_x_log10(limits = c(0.5, 50), breaks = c(0.5, 1, 2, 5, 10, 20, 50)) + 
+    scale_x_continuous(limits = c(0.5, 30), breaks = c(0.5, 1, 2, 5, 10, 20)) + 
     scale_y_continuous(limits = c(-0.25, 0.25) ) +
@@
     distinct() %>% 
-    select(dz_2011, distance_to_centre, year, pop_incomedeprived) %>% 
+    dplyr::select(dz_2011, distance_to_centre, year, pop_incomedeprived) %>% 
     mutate(distance_to_centre = distance_to_centre / 1000) %>% 
     spread(year, pop_incomedeprived) %>% 
-    mutate(change = (`2012` / `2004`) - 1) %>% 
+    mutate(change = (`2016` / `2004`) - 1) %>% 
     ggplot(., aes(x = distance_to_centre, y = change)) + 
@@
     distinct() %>% 
-    select(dz_2011, distance_to_centre, year, pop_total) %>% 
+    dplyr::select(dz_2011, distance_to_centre, year, pop_total) %>% 
     mutate(distance_to_centre = distance_to_centre / 1000) %>% 
     spread(year, pop_total) %>% 
-    mutate(change = (`2012` / `2004`) - 1) %>% 
+    mutate(change = (`2016` / `2004`) - 1) %>% 
     ggplot(., aes(x = distance_to_centre, y = change)) + 
@@
 lkup  %>%
-  select(dz_2001, dz_2011)  %>%
+  dplyr::select(dz_2001, dz_2011)  %>%
   arrange(dz_2001, dz_2011)  %>%
@@
   arrange(dz_2011, dz_2001)  %>%
-  select(dz_2011, dz_2001, n)  %>%
+  dplyr::select(dz_2011, dz_2001, n)  %>%
   group_by(dz_2011)   %>%
@@
   left_join(simd_combined, by = c("dz_2001" = "datazone"))  %>%
-  select(-simd_rank)  %>%  # Not meaningful to reweight rank
+  dplyr::select(-simd_rank)  %>%  # Not meaningful to reweight rank
   group_by(dz_2011, year) %>%

### 2016-08-31 scripts/combine_simds.R

@@ -13,6 +13,17 @@ pacman::p_load(
 
 # Load SIMDs --------------------------------------------------------------
 
+# income deprived part
+simd_2016_id <- read_excel(path = "data/simd/00504760.xlsx", sheet = "Data", col_types = rep("text", 36))
+# SIMD 2016
+# 00504760.xlsx
+simd_2016_id[] <- lapply(simd_2016_id, function(x) str_replace(x, "\\*", 0))
+simd_2016_id[,4:36] <- lapply(simd_2016_id[,4:36], as.numeric)
+
+# For SIMD ranks (scores not published!)
+simd_2016 <- read_excel(path = "data/simd/00504759.xlsx", sheet = "SIMD16 ranks")
+
+
 simd_2012 <- read_excel(path = "data/simd/00410767.xls")
 # SIMD 2012
 # 00410767.xls
@@ -84,12 +95,41 @@ simd_2012 %>%
   mutate(year = 2012) %>% 
   select(datazone, year, everything()) -> simd_2012_simple
 
+simd_2016_id %>% 
+  select(
+    datazone = Data_Zone, 
+    pop_total = Total_population, 
+    pop_workingage =Working_age_population, 
+    pop_incomedeprived = `Income_count`
+  ) %>% 
+  mutate(year = 2016) %>% 
+  select(datazone, year, everything()) -> simd_2016_simple_id
+
+simd_2016 %>% 
+  select(
+    datazone = Data_Zone,
+    simd_rank = Overall_SIMD16_rank
+  ) %>% 
+  mutate(
+    simd_score = NA
+  ) %>% 
+  inner_join(simd_2016_simple_id) %>% 
+  select(
+    datazone, year, 
+    pop_total, pop_workingage, 
+    pop_incomedeprived, simd_score, simd_rank
+  ) -> simd_2016_simple
+rm(simd_2016_simple_id)
+
+
+
 # combine these 
 simd_combined <- bind_rows(
   simd_2004_simple, 
   simd_2006_simple, 
   simd_2009_simple, 
-  simd_2012_simple
+  simd_2012_simple,
+  simd_2016_simple
 ) %>% 
   filter(!is.na(datazone))
 

### 2016-08-31 scripts/do_dz2011_data_management.R

@@ -5,23 +5,30 @@ simd_combined <- read_csv("data/simd/simd_combined.csv")
 # Paul Norman Lookup
 lkup <-  read_csv("data/paul_norman_file/paul_norman_dz2011_table.csv")
 
+# simd 2016 already dz2011, so exclude from reweighting below
 
+simd_older_combined <- simd_combined %>% filter(year != 2016)
 # Produce reweighted SIMD scores
 lkup  %>% 
-  select(dz_2001, dz_2011)  %>% 
+  dplyr::select(dz_2001, dz_2011)  %>% 
   arrange(dz_2001, dz_2011)  %>%  
   group_by(dz_2001, dz_2011)  %>% 
   tally  %>% # produce n, giving number of OAs which contain particular groupings of dz_2001 and dz_2011
   arrange(dz_2011, dz_2001)  %>% 
-  select(dz_2011, dz_2001, n)  %>% 
+  dplyr::select(dz_2011, dz_2001, n)  %>% 
   group_by(dz_2011)   %>% 
   arrange(dz_2011, dz_2001) %>% 
   mutate(weight = n / sum(n))   %>% # Weighting by dz_2011
-  left_join(simd_combined, by = c("dz_2001" = "datazone"))  %>% 
-  select(-simd_rank)  %>%  # Not meaningful to reweight rank
+  left_join(simd_older_combined, by = c("dz_2001" = "datazone"))  %>% 
+  dplyr::select(-simd_rank)  %>%  # Not meaningful to reweight rank
   group_by(dz_2011, year) %>% 
   summarise_each( ~ sum(. * weight, na.rm =T), 6:9) -> simd_2011_reweighted
 
+simd_2016_simplified <- simd_combined %>% 
+  filter(year == 2016) %>% 
+  dplyr::select(dz_2011 = datazone, year, pop_total, pop_workingage, pop_incomedeprived, simd_score)
+
+simd_2011_reweighted <- bind_rows(simd_2011_reweighted, simd_2016_simplified)
 
 # Shapefile for 2011
 dz_2011 <- read_shape(file = "shapefiles/SG_DataZoneBdry_2011/SG_DataZone_Bdry_2011.shp")

### 2016-08-31 scripts/from_gavin/do_rci_and_d_for_all_permutations.R

@@ -37,15 +37,15 @@ do_model <- function(place, initial_dist = 18000){
   }
   
   simd_2011_reweighted %>% 
-    select(dz_2011, year, pop_total, pop_incomedeprived) -> tmp
+    dplyr::select(dz_2011, year, pop_total, pop_incomedeprived) -> tmp
   
   tmp %>% 
-    select(-pop_incomedeprived) %>% 
+    dplyr::select(-pop_incomedeprived) %>% 
     mutate(year = paste0("pop_total_", year)) %>% 
     spread(year, pop_total) -> pops
   
   tmp %>% 
-    select(-pop_total) %>% 
+    dplyr::select(-pop_total) %>% 
     mutate(year = paste0("pop_id_", year)) %>% 
     spread(year, pop_incomedeprived) -> incdeps
   
@@ -78,7 +78,7 @@ do_model <- function(place, initial_dist = 18000){
     as.integer(dz_city@data$pop_id_2006), 
     as.integer(dz_city@data$pop_id_2009), 
     as.integer(dz_city@data$pop_id_2012),
-    as.integer(dz_city@data$pop_id_2012) # placeholder for new data
+    as.integer(dz_city@data$pop_id_2016) 
   )
   
   N.mat <- cbind(
@@ -86,8 +86,13 @@ do_model <- function(place, initial_dist = 18000){
     as.integer(dz_city@data$pop_total_2006), 
     as.integer(dz_city@data$pop_total_2009), 
     as.integer(dz_city@data$pop_total_2012),
-    as.integer(dz_city@data$pop_total_2012) # placeholder for new data
+    as.integer(dz_city@data$pop_total_2016) 
   )
+  
+  denom_too_small <- N.mat[,5] < 1
+  N.mat[denom_too_small, 5] <- 1
+  
+  Y.mat[denom_too_small, 5] <- 0
 
 
   
@@ -116,15 +121,15 @@ do_model <- function(place, initial_dist = 18000){
   #### Compute the global RCI and D
   indicators.post <- array(NA, c(n.keep,10))
   colnames(indicators.post) <- c(
-    "RCI_2004", "RCI_2006","RCI_2009","RCI_2012", "RCI_2015", 
-    "D_2004", "D_2006", "D_2009", "D_2012", "D_2015"
+    "RCI_2004", "RCI_2006","RCI_2009","RCI_2012", "RCI_2016", 
[…]

### 2016-08-31 markdown/draft.RmD
rerun and upload markdown

@@
 date: "31 August 2016"
-output: word_document
+output: html_document
 ---
@@
 
-* [CHECK] In all four of Scotland’s major cities, the number of poor or ‘income deprived’ people in inner city areas fell between 2012 and 2016 according to the latest Scottish Index of Multiple Deprivation data, while the number of non-poor people in these areas rose.
-*	[CHECK] In less central areas, the population of both groups was little changed. Poverty therefore became more ‘suburbanised’ in this period.
-*	[CHECK] This continues the trend observed over the last four waves of the SIMD, from 2004 to 2012, again across all four cities. 
-*	[CHECK] The total population in the central areas rose in all four cities between 2004 and 2016.
-This is generally to be welcomed from an environmental point of view.
-It suggests that planning policies have succeeded in encouraging people to live in higher density locations which tend to be better served by public transport.
-*	[CHECK] However the displacement of lower income groups suggests these policies will fail to deliver the socially mixed communities which planners have so frequently advocated as part of their vision for sustainable cities.
-*	[CHECK] The main reason for this change is likely to be the continued reductions in housing subsidies for lower income households.
+* In all four of Scotland’s major cities, poverty has become more suburbanized over the last 12 years.
+The proportion of poor or ‘income deprived’ people in inner city areas fell, while the number of non-poor people in these areas rose.
+*	The latest data, published as part of the Scottish Index of Multiple Deprivation today, confirm this trend, albeit with some variation between cities.
+The process is advancing most in Edinburgh with Glasgow second.
+It is slowest in Dundee.
+*	The central areas of all four cities retained their share of the total city population or saw this rise between 2004 and 2016.
+This is a significant change from decades of population losses and is generally to be welcomed from an environmental point of view.
+It suggests that planning policies have succeeded in encouraging people to live in higher-density locations which tend to be better served by public transport, reducing reliance on cars.
+*	However the displacement of lower income groups suggests these policies will fail to deliver the socially mixed communities which planners have so frequently advocated as part of their vision for sustainable cities.
+*	The main reason for this change is likely to be the continued reductions in housing subsidies for lower income households.
 The last 20 years have seen the loss of social rented housing and greater reliance on private renting, combined with the reductions in housing benefits.
@@
 They have sought to limit suburbanization and direct new housing development towards brownfield land.
-In many cases, these policies were driven by a vision of the “sustainable city”, based around higher densities, effective public transport and
- hoped these changes would bring new opportunities for more deprived communities as well as improvements in the physical environment.
-They sought to grow the inner city population by developing vacant or underused land, producing mixed communities.
+In many cases, these policies were driven by a vision of the “sustainable city”, based around higher densities, effective public transport and socially mixed communities.
+Governments hoped these changes would bring new opportunities for more deprived inner city areas as well as improvements in the physical environment.
+They sought to grow the inner city population by developing vacant or underused land.
 
@@
 
-These analyses were expanded to the four major Scottish cities - Aberdeen, Dundee, Edinburgh and Glasgow - for a paper to appear in Annals of the American Association of Geographers, by the same authors.
+These analyses were expanded to the four major Scottish cities - Aberdeen, Dundee, Edinburgh and Glasgow - for a paper to appear shortly in *Annals of the American Association of Geographers*, by the same authors.
 This paper calculated 95% credible intervals (CrIs) for the difference in RCI from 2001 to 2011 for the same three measures of poverty, and for each of the four cities.
@@
 Different editions of the SIMD have been produced in 2004, 2006, 2009 and 2012, with the latest release on 31 August 2016.
+
 Data are made available for the small area units known as Datazones, equivalent to the Lower Super Output Area (LSOA) in England & Wales.
@@
 For 2016, the SIMD moved over to using the new (2011) Datazone boundaries.
+
 The methodology for constructing the SIMD has been consistent over this period although details differ slightly from one edition to the next.
@@
 One year (2006) is perhaps slightly different due to the omission of households in receipt of tax credits on that occasion so results for that year should perhaps be treated with more caution.
+
 In a recent assessment, Bailey et al (2016) have shown that there is no evidence of any urban-rural bias in this measure.
+
 SIMD data releases for 2004, 2006, 2009, 2012 and 2016 were downloaded from the SIMD website.
@@
 
-The following plot shows scatterplots of the ID proportions in each Scottish datazone in 2004, 2009 and 2012.
+The following plot shows scatterplots of the ID proportions in each Scottish datazone in 2004 through to 2016.
 Within these plots the blue line indicates parity between values on the x and y axis, and the red dashed line shows a nonlinear smoother function between the two ID proportions being plotted, and hence the level of deviation between the actual relationship over the two years and parity.
@@
 
-The first of the plots, 2004 compared with 2012, shows strong spatial dependence in areas with low to moderate ID proportions, but much less temporal dependence in areas with high ID proportions in 2004.
+The first of the plots, 2004 compared with 2016, shows strong spatial dependence in areas with low to moderate ID proportions, but less temporal dependence in areas with high ID proportions in 2004.
 The second subfigure plots 2004 against 2006 values, and shows a near linear temporal dependence for the complete range of 2004 values, with a global trend towards lower ID proportions throughout Scotland.
-The temporal relationship between 2006 and 2009 starts to become nonlinear at the upper range of 2006 values, and this nonlinearity at the upper range becomes much stronger between 2009 and 2012.
+The temporal relationship between 2006 and 2009 starts to become nonlinear at the upper range of 2006 values, and this nonlinearity at the upper range becomes much stronger between 2009 and 2012, and 2012 and 2016.
 
-These plots indicate a complex change in the temporal dependence of ID proportions, likely occurring over the period 2008 to 2012.
-SIMD 2016 values will be invaluable in indicating whether these changes have continued, altered or settled between 2012 and 2016.
+There is no suggestion that the change between 2012 and 2016 is significantly different to that for the earlier periods, giving reassurance that the process of moving the earlier data to the new boundaries is not distorting the picture.
 
@@
 
-Below is a table showing the percentage if the population in each city (within 15 km of the centre) defined as 'income deprived' according to SIMD.
+Below is a table showing the percentage if the population in each city (within 20 km of the centre) defined as 'income deprived' according to SIMD.
 In 2004 these percentages ranged from 9.8% in Aberdeen to 20.8% in Glasgow.
-By 2012 they had fallen to 7.9% in Aberdeen to 17.1% om Glasgow.
-In every city except Aberdeen the percentage increased from 2006 to 2009, as would be expected given the onset of the 2008 recession within this period, but then fell again from the 2009 to the 2012 SIMD release.
+By 2012 they had fallen to 7.9% in Aberdeen to 17.1% in Glasgow.
+In every city except Aberdeen the percentage increased from 2006 to 2009, as would be expected given the onset of the 2008 recession within this period, but then fell again from the 2009 to the 2012 and 2016 SIMD releases.
 
@@
 
-[CHECK] The overall trend for all four cities is for poverty to decentralize between 2004 to 2016.
+The overall trend for all four cities is for poverty to decentralize between 2004 to 2016.
 Dundee shows the most centralisation of poverty, then Aberdeen, and Glasgow, with Edinburgh the most decentralised.
-Indeed, from 2009 onwards, Edinburgh is unique in having poorer households living further from the city centre than non-poor.
+Indeed, from 2009 onwards, Edinburgh is unique in having poorer households living further from the city centre than non-poor on average, and this increased in the most recent period.
 
-[CHECK] For Glasgow and Dundee, the decline in RCI is continuous, whereas for Aberdeen and Edinbugh RCI values increased from 2004 to 2006, before declining in 2009 and 2012.
-As note above, there is some reason to be more cautious about the comparability of the 2006 data with that from other years.
+For Glasgow and Dundee, the decline in RCI is continuous although, in the most recent period, the change is within the limits of random fluctuation.
+For Aberdeen and Edinburgh, RCI values increased slightly from 2004 to 2006, before declining after 2009.
+As noted above, there is some reason to be more cautious about the comparability of the 2006 data with that from other years.
 
@@
 
-As with RCI, Dissimilarity scores have also fallen over time for all four cities.
+As with RCI, Dissimilarity scores have also fallen over time for all four cities although Edinburgh shows signs of an increase in the most recent period.
 The greatest contrast in the trends is between Aberdeen and Edinburgh.
-In 2004 both cities had similar and overlapping D scores, but between 2004 and 2009 segregation fell rapidly for Edinburgh, while remaining fairly stable in Aberdeen, before falling from 2009 to 2012.
+In 2004 both cities had similar and overlapping D scores, but between 2004 and 2009 segregation fell rapidly for Edinburgh, while remaining fairly stable in Aberdeen, before falling from 2009 to 2016.
 Glasgow has always had the lowest D scores of the four cities, and Dundee the highest, but since 2009 the distribution of D estimates for Edinburgh have started to overlap those of Glasgow as the rate of decline in D has been slightly greater in Edinburgh than Glasgow.
@@
 
-[CHECK] In the figure below, the vertical axis shows the change in the proportion income deprived from 2004 to 2016, and the horizontal axis shows the distance of datazones from the city centres, plotted on a log scale, for each of the four main Scottish cities.
+In the figure below, the vertical axis shows the change in the proportion of people income deprived from 2004 to 2016, and the horizontal axis shows the distance of datazones from the city centres, plotted on a log scale, for each of the four main Scottish cities.
 Individual datazones are indicated with faint circles, and a nonlinear smoother has been added showing how this the change in income deprivation varies as a function of distance.
@@
 
-These plots suggest a qualitatively similar relationship between change in deprivation and distance, up to around 12km from city centre, for Aberdeen, Edinburgh and Glasgow, with Glasgow showing the strongest gradient of change between these two periods.
-Dundee shows a similar kind of relationship, but concentrated within the first 5km, rather than 12km, of distance from city centre.
-This is consistent with the maps of income deprivation in Dundee, showing much greater levels of spatial concentration than for the other cities.
+In broad terms, these plots suggest a similar picture, with reductions in deprivataion rates greater in areas closest to the centre.
 
@@
 changes to total population sizes.
-To gain a better understanding of the relative influence of these two changes to the relationship shown in the previous figure, the next two figures show the proportionate change from 2004 to 2012 in income deprived populations, and total populations.
+To gain a better understanding of the relative influence of these two changes to the relationship shown in the previous figure, the next groups of figures show the proportionate change from 2004 to 2016 in income deprived populations, and total populations.
+These plots indicate greater proportionate falls in the ID population shares at distances closes to the city centre.
 
@@
 ```
-This plot indicates proportionate falls in the ID population shares at distances closes to the city centre, up to around 20% in Aberdeen and Glasgow, and either proportionate rises (Dundee and Edinburgh) or little change in proportions (Aberdeen and Glasgow) at greater distance from the city centre.
-This seems to suggest, for Aberdeen and Glasgow in particular, that 'gentrification of the centre' could be as much or more of an explanation for falls in RCI as 'impoverishment of the periphery', though that there are somewhat different socioeconomic mechanisms in place for the four cities.
+The plots suggest that more central areas have held or increased their share of the total city population.
+Changes have been the same or slightly greater in Datazones closer to the centre.
+In Edinburgh, the centre has grown at the expense of surrounding areas.
+In Glasgow, it has declined less rapidly.
 

### 2016-08-31 Script.R
add figs for conversation

@@ -21,6 +21,58 @@ source('scripts/from_gavin/binomial.MCARleroux.R')
 Rcpp::sourceCpp('scripts/from_gavin/aqmen.cpp')
 
 
+
+
+# Figure for The Conversation
+
+simd_2011_reweighted  %>% 
+  ungroup() %>%  
+  left_join(centre_distance, by = c("dz_2011" = "datazone")) %>% 
+  distinct() %>%
+  rename(City = nearest_centre) %>% 
+  filter(City %in% c("Aberdeen", "Dundee", "Edinburgh", "Glasgow")) %>% 
+  mutate(
+    lower_thresh = ifelse(City %in% c("Aberdeen", "Dundee"), 3200, 5000),
+    centralness = ifelse(distance_to_centre < 20000, "periphery", "exurb"),
+    centralness = ifelse(distance_to_centre < lower_thresh, "centre", centralness)
+  ) %>% dplyr::select(-lower_thresh) %>% 
+  filter(centralness %in% c("centre", "periphery")) %>% 
+  group_by(City, year, centralness) %>%
+  mutate(pop_nonid = pop_total - pop_incomedeprived) %>% 
+  summarise(pop_nonid = sum(pop_nonid), pop_incomedeprived = sum(pop_incomedeprived)) %>% 
+  summarise(
+    `Income Deprived` = pop_incomedeprived[centralness == "centre"] / sum(pop_incomedeprived),
+    `Not Income Deprived` = pop_nonid[centralness == "centre"] / sum(pop_nonid)
+  ) %>% 
+  gather(key = "type", value = "prop", `Income Deprived`:`Not Income Deprived`) -> tmp
+
+
+tmp %>% filter(type == "Income Deprived") %>% 
+  mutate(prop = 100 * prop) %>% 
+  ggplot(., 
+         aes(
+           x = year, y = prop 
+         )
+  ) + 
+  geom_line() + geom_point() +
+  facet_wrap( ~ City, nrow = 2, scales = "free_y") + 
+  labs(x = "Year", y = "Percentage", title = "Percent income deprived near City Centre") 
+
+ggsave("figures/conversation_01.png", height = 15, width = 15, units = "cm", dpi = 300)
+
+
+tmp %>% filter(type == "Not Income Deprived") %>% 
+  mutate(prop = 100 * prop) %>% 
+  ggplot(., 
+         aes(
+           x = year, y = prop 
+         )
+  ) + 
+  geom_line() + geom_point() +
[…]

### 2016-10-28 markdown/draft.RmD
make slight edit to markdown draft

@@
 The proportion of poor or ‘income deprived’ people in inner city areas fell, while the number of non-poor people in these areas rose.
-*	The latest data, published as part of the Scottish Index of Multiple Deprivation today, confirm this trend, albeit with some variation between cities.
+*	The latest data, published as part of the Scottish Index of Multiple Deprivation (SIMD) on 31 August 2016, confirm this trend, albeit with some variation between cities.
 The process is advancing most in Edinburgh with Glasgow second.

### 2016-10-28 markdown/draft_as_notebook.Rmd
start on presentation; create notebook doc

--- 2016-10-28 markdown/draft.RmD
@@
 date: "31 August 2016"
-output: html_document
+output: html_notebook
 ---
@@
 
-```{r show_id_map, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE}
+```{r show_id_map, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE, dpi = 300, fig.width = 15}
 simd_2011_reweighted %>% 
@@
 
-```{r show_id_map_glasgow, echo = F, eval = T, message = FALSE, warning = FALSE}
+```{r show_id_map_glasgow, echo = F, eval = T, message = FALSE, warning = FALSE, dpi = 300, fig.width = 15}
 
@@
 
-```{r show_id_map_edinburgh, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE}
+```{r show_id_map_edinburgh, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE, dpi = 300, fig.width = 15}
 
@@
 
-```{r show_id_map_dundee, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE}
+```{r show_id_map_dundee, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE, dpi = 300, fig.width = 15}
 
@@
 
-```{r show_id_map_aberdeen, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE}
+```{r show_id_map_aberdeen, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE, dpi = 300, fig.width = 15}
 
@@
 
-```{r show_temporal_dep, echo = FALSE, warning = FALSE, EVAL = TRUE}
+```{r show_temporal_dep, echo = FALSE, warning = FALSE, EVAL = TRUE, dpi = 300, fig.width = 15}
 simd_2011_reweighted %>% 
@@
     arrange(`2004`) %>% 
-    kable(., format = "markdown", caption = "Income deprived % by City and Year")
+    kable(., format = "markdown", caption = "Income deprived % by City and Year", )
 
@@
 
-```{r plot_change_rci, eval = TRUE, echo = FALSE, warning = FALSE}
+```{r plot_change_rci, eval = TRUE, echo = FALSE, warning = FALSE, dpi = 300}
 # Do RCI graph
@@
 
-```{r plot_change_d, eval = TRUE, echo = FALSE, warning = FALSE}
+```{r plot_change_d, eval = TRUE, echo = FALSE, warning = FALSE, dpi = 300}
 # Do D graph
@@
 
-```{r show_change_in_id, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+```{r show_change_in_id, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE, dpi = 300, fig.width = 20, fig.height = 15}
 decent_plot_for_city <- function(this_city){
@@
 
-```{r plot_changeidpop, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+```{r plot_changeidpop, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE,dpi = 300, fig.width = 20, fig.height = 15}
 idpopchange_plot_for_city <- function(this_city){
@@
 
-```{r plot_changetotalpop, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+```{r plot_changetotalpop, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE, dpi = 300, fig.width = 20, fig.height = 15}
 totalpopchange_plot_for_city <- function(this_city){

### 2016-11-04 markdown/cart.Rmd
look at cart

---
title: "CART and Cluster analysis"
output: html_notebook
---

This is a short attempt to use SIMD data to demonstrate CARTs, Cluster Analyses etc.

```{r setup01, include=TRUE, echo = FALSE, message = FALSE, warning = FALSE}
rm(list = ls())
# Code chunk example from http://stackoverflow.com/questions/24585254/working-with-knitr-using-subdirectories
library(knitr)

opts_knit$set(root.dir=normalizePath('../'))
```

```{r setup02, include = TRUE, cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE}

require(pacman)

pacman::p_load(
  readr, readxl,
  purrr, stringr, tidyr, dplyr,
  broom,
  rgeos,
  ggplot2, cowplot,
  tmap, RColorBrewer,

  rpart, rpart.plot
)

simd_2016 <- read_excel(path = "../data/simd/00504759.xlsx", sheet = "SIMD16 ranks")

```

# CART

Classification and regression trees try to classify observations into discrete groups based on the characteristics of variables. They are predictive rather than descriptive, unlike cluster analysis, as the discrete groups are known beforehand, and predicted can be compared against actual groups. Through this they can be helpful for getting a sense of the main features that distinguish different groups, and so perhaps the discriminatory power (statistical usefulness) of different variables available to you. 
In this example they will try to classify datazones into council areas based on various 2016 SIMD ranks, and population sizes.

```{r cart_example}
# Using examples from 
#http://www.statmethods.net/advstats/cart.html

cart_model <- rpart(Council_area ~ Total_population + Working_age_population + Overall_SIMD16_rank + Income_domain_2016_rank + Housing_domain_2016_rank + Access_domain_2016_rank, data = simd_2016, method = "class")

```

```{r plot_tree}
# plot tree

prp(cart_model, uniform=F, 
  	main="Classification Tree for LAs based on SIMD characteristics",
    clip.right.labs = F, varlen = 0)
#text(cart_model, use.n=F, all=F)

```

This is suggesting that the housing domain rank is the most useful single indicator for identifying one c local authority area, namely Glasgow city. If the rank in the housing domain is below 1134, then it predicts the datazone is in Glasgow City.
Then, it suggests that if an areas has an overall SIMD rank above 6536 (there are now 6976 2011 datazones, up from 6505 datazones based on the 2001 census), then predict City of Edinburgh. Then, if the access domain rank is less than 602, predict Aberdeenshire, and so on.

```{r further_stats}

printcp(cart_model) # display the results 
plotcp(cart_model) # visualize cross-validation results 
summary(cart_model) # detailed summary of splits

```

### 2017-02-13 scripts/calc_rci_by_place_imd.R
redo all w/ Scotland Cities

@@ -1,5 +1,9 @@
 # Explore IMD by city 
 
+
+# TO DO: Wales  IMD 
+
+
 rm(list = ls())
 
 require(pacman)
@@ -14,11 +18,6 @@ pacman::p_load(
 )
 
 source("scripts/from_gavin/RCI.R")
-# imd
-dist_to_centre <- read_csv("data/lsoa_2011_by_dist_to_centres.csv")
-dta <- read_csv("data/imd/imd_id_tidied.csv", col_types = "icdd")
-
-
 
 
 D <- function(minority, total){
@@ -34,6 +33,16 @@ D <- function(minority, total){
   out <- 0.5 * sum(ad)
   out
 }
+
+
+
+# imd
+dist_to_centre <- read_csv("data/lsoa_2011_by_dist_to_centres.csv")
+dta <- read_csv("data/imd/imd_id_tidied.csv", col_types = "icdd")
+
+#SIMD
+dist_to_centre_scot <- read_csv("data/dz_2011_by_dist_to_centres.csv")
+dta_scot <- read_csv("data/simd/simd_combined_on_2011.csv")
 
 # Inputs to RCI are 
 #povvec
@@ -65,15 +74,59 @@ dta %>%
   select(place, year, rci = rci_val, rdi = rdi_val, d = d_val) %>% 
   ungroup -> rci_by_year_place
 
+
+# Same but for Scotland
+dta_scot %>% 
+  inner_join(dist_to_centre_scot, by = c("dz_2011" = "dz")) %>%
+  mutate(pdens = pop_total / area) %>%
+  select(dz_2011, place, year, pop_id = pop_incomedeprived, pop_total, distance, pdens) %>% 
+  group_by(place, year) %>% 
+  nest() %>% 
+  mutate(
+    pvec = map(data, ~ .[["pop_id"]]),
+    tvec = map(data, ~ .[["pop_total"]]),
+    ordr = map(data, ~ order(.[["distance"]])),
+    ordr_dens = map(data, ~ order(.[["pdens"]], decreasing = T))
+  ) %>% 
+  mutate(rci_val = pmap_dbl(list(pvec, tvec, ordr), RCI)) %>% 
+  mutate(rdi_val = pmap_dbl(list(pvec, tvec, ordr_dens), RCI)) %>% 
+  mutate(d_val = pmap_dbl(list(pvec, tvec), D)) %>% 
+  select(place, year, rci = rci_val, rdi = rdi_val, d = d_val) %>% 
[…]

### 2017-02-13 scripts/do_dz2011_data_management.R

@@ -30,6 +30,8 @@ simd_2016_simplified <- simd_combined %>%
 
 simd_2011_reweighted <- bind_rows(simd_2011_reweighted, simd_2016_simplified)
 
+write_csv(x = simd_2011_reweighted, path = "data/simd/simd_combined_on_2011.csv")
+
 # Shapefile for 2011
 dz_2011 <- read_shape(file = "shapefiles/SG_DataZoneBdry_2011/SG_DataZone_Bdry_2011.shp")
 

### 2017-02-13 scripts/imd_by_ttwa.R

@@ -14,6 +14,19 @@ pacman::p_load(
 )
 
 
+# England and Wales  ------------------------------------------------------
+
+
+# IMD first, then SIMD 
+# # Shapefile for 2011
+lsoa_2011 <- read_shape(file = "shapefiles/Lower_layer_super_output_areas_(E+W)_2011_Boundaries_(Full_Clipped)_V2/LSOA_2011_EW_BFC_V2.shp")
+
+
+# 
+
+# Non Shapefile data ------------------------------------------------------
+
+
 # imd
 dta <- read_csv("data/imd/imd_id_tidied.csv", col_types = "icdd")
 
@@ -58,11 +71,13 @@ ttwas_to_keep <- c(
   Southend = "E30000268",
   Oxford = "E30000250",
   Portsmouth = "E30000254",
-  Reading = "E30000256"
-  
+  Reading = "E30000256",
+  # Wales
+  Cardiff = "W22000024"
+  # Scotland
+#  Glasgow = "S22000065",
+#  Edinburgh = "S22000059"
 )
-
-
 
 lsoa_id_by_ttwa %>% 
   filter(ttwa %in% ttwas_to_keep) %>% 
@@ -87,8 +102,6 @@ ttwa_centres <- c(
   Sheffield   = "E01033264",
   Leeds       = "E01033010",
   Nottingham  = "E01033406",
-  
-  
   `Warrington and Wigan` = "E01033300",
   `Wolverhampton and Walsall` = "E01010521",
   Luton = "E01015794",
@@ -101,17 +114,20 @@ ttwa_centres <- c(
   Southend = "E01015852",
   Oxford = "E01028549",
   Portsmouth = "E01017028",
-  Reading = "E01016358"
-  
-)
+  Reading = "E01016358",
+  # Scottish Cities
+  # Aberdeen = "S01006646",
+  # Glasgow = "S01010265",
+  # Edinburgh = "S01008677",
+  # Dundee = "S01007705",
+  # Inverness = "S01010620",
+  # Perth = "S01011939",
+  # `Falkirk and Stirling` = "S01013067",
+  # Cardiff
+  Cardiff = "W01001939" #Cardiff 03FF
+)            
 
-# Shapefile for 2011
-lsoa_2011 <- read_shape(file = "shapefiles/Lower_layer_super_output_areas_(E+W)_2011_Boundaries_(Full_Clipped)_V2/LSOA_2011_EW_BFC_V2.shp")
 
[…]

### 2017-02-14 scripts/stats_by_ttwa.Rmd
save figures in figures/ttwa

---
title: "RCI, RDI and D for English and Scottish Cities"
author: "Jon Minton"
date: "14 February 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2017-02-14 scripts/calc_rci_by_place_imd.R

@@ -9,8 +9,9 @@ rm(list = ls())
 require(pacman)
 
 pacman::p_load(
+  tidyverse,
   readr, readxl,
-  stringr, tidyr, dplyr, 
+  stringr,  
   purrr,
   rgeos,
   ggplot2,
@@ -136,14 +137,20 @@ rci_by_year_place %>%
   ggplot(., aes(x = year, y = rci)) + 
   geom_line() + geom_point() + 
   geom_hline(aes(yintercept = 0)) +  
-  facet_wrap(~place)
+  facet_wrap(~place) + 
+  labs(title = "RCI by year and TTWA", subtitle = "TTWAs arranged by size", x = "Year", y = "Relative Centralisation Index (RCI)")
+ggsave("figures/TTWA/rci_by_year_ttwa.png", height= 25, width = 25, units = "cm", dpi = 300)
+
 
 rci_by_year_place %>% 
   mutate(place = factor(place, ordered = T, levels = place_by_size_order)) %>% 
   ggplot(., aes(x = year, y = rdi)) + 
   geom_line() + geom_point() + 
   geom_hline(aes(yintercept = 0)) +  
-  facet_wrap(~place)
+  facet_wrap(~place) + 
+  labs(title = "RDI by year and TTWA", subtitle = "TTWAs arranged by size", x = "Year", y = "Relative 'Densification' Index (RDI)")
+ggsave("figures/TTWA/rdi_by_year_ttwa.png", height= 25, width = 25, units = "cm", dpi = 300)
+
 
 # RDI and RCI on same plot 
 rci_by_year_place %>% 
@@ -154,7 +161,10 @@ rci_by_year_place %>%
   geom_line(aes(linetype = measure)) + 
   geom_point(aes(shape = measure)) +
   geom_hline(aes(yintercept = 0)) + 
-  facet_wrap(~place)
+  facet_wrap(~place) +
+  labs(title = "RDI and RCI by year and TTWA", subtitle = "TTWAs arranged by size", x = "Year", y = "Score")
+ggsave("figures/TTWA/rdi_rci_by_year_ttwa.png", height= 25, width = 25, units = "cm", dpi = 300)
+
 
 
 
@@ -167,86 +177,24 @@ rci_by_year_place %>%
   geom_hline(yintercept = 0) + 
   coord_cartesian(ylim = c(0.25, 0.45)) + 
   labs(x = "Year", y = "Dissimilarity Index") + 
[…]

### 2017-02-14 scripts/combine_wimds.R

@@ -1,0 +1,362 @@
+rm(list = ls())
+
+require(pacman)
+
+pacman::p_load(
+  readr, readxl,
+  stringr, tidyr, dplyr, 
+  rgeos,
+  ggplot2,
+  tmap
+)
+
+
+
+# Merging of WIMD data ----------------------------------------------------
+
+
+# NOTE: Appropriate data have not yet been identified from the Welsh website.
+# A request was sent for appropriate data on 14/2/2017
+
+
+# 2015 Data  --------------------------------------------------------------
+
+#The aim of the data coding is to estimate the number of people income deprived in the census. This 
+# appears to require combining data from file 5 and file 6. 
+
+# Let's explore further... 
+
+imd_num <- read_excel("data/imd/2015/File_5_ID_2015_Scores_for_the_Indices_of_Deprivation.xlsx", "ID2015 Scores")
+imd_num %>% 
+  .[,c(1, 6)] -> imd_num
+
+names(imd_num) <- c("lsoa", "prop_id")
+
+# Now the denominator
+
+imd_denom <- read_excel("data/imd/2015/File_6_ID_2015_Population_denominators.xlsx", 
+                        sheet = 2, col_types = rep("text", 9), skip = 1,
+                        col_names = c(
+                          "lsoa", "lsoa_nm", "lad", "lad_nm", "pop_total", "pop_children", "pop_wage", "pop_older", "pop_wage2"
+                        )
+)
+
+imd_denom %>% 
+  select(lsoa, pop_total) -> imd_denom
+
+imd_denom %>% 
+  mutate(pop_total = as.numeric(str_replace_all(pop_total, ",", ""))
+  ) -> imd_denom
+
+
+imd_2015 <- imd_num %>% inner_join(imd_denom)
+
+imd_2015 %>% 
+  mutate(pop_id = prop_id * pop_total) %>% 
+  mutate(year = 2015) %>% 
+  select(lsoa, year, pop_id, pop_total) -> imd_2015
+
+
+# 
+
+# 2010 income deprived ----------------------------------------------------
+
+
+# The two fields to use are 
[…]

### 2017-02-16 markdown/3D_plots.Rmd
add concentration scores

---
title: "3D relationships between RDI, RCI and D in TTWAs"
author: "Jon Minton"
date: "15 February 2017"
output: html_document
---

```{r setup, include=FALSE}
rm(list = ls())
# Code chunk example from http://stackoverflow.com/questions/24585254/working-with-knitr-using-subdirectories
library(knitr)
library(rgl)
opts_knit$set(root.dir=normalizePath('../'))
knit_hooks$set(webgl = hook_webgl)

```

# Aim

This document will (hopefully) allow D, RCI and RDI to be explored interactively for each of the TTWAs. 
For each TTWA, the first period is indicated with a blue sphere, the last with a red sphere, and intermediate spheres along the path are grey.

```{r setup2, include = F, echo = F}

require(pacman)

pacman::p_load(
  tidyverse,
  readr, readxl,
  forcats,
  stringr,  
  purrr,
  rgeos,
  ggplot2,
  tmap
)

source("scripts/from_gavin/RCI.R")

D <- function(minority, total){
  majority      <- total - minority
  MINORITY      <- sum(minority)
  MAJORITY      <- sum(majority)

  p1 <- minority / MINORITY
  p2 <- majority / MAJORITY

  ad <- abs(p1 - p2)

  out <- 0.5 * sum(ad)
  out
}

# imd
dist_to_centre <- read_csv("data/lsoa_2011_by_dist_to_centres.csv")
dta <- read_csv("data/imd/imd_id_tidied.csv", col_types = "icdd")

#SIMD
dist_to_centre_scot <- read_csv("data/dz_2011_by_dist_to_centres.csv")
dta_scot <- read_csv("data/simd/simd_combined_on_2011.csv")

# Inputs to RCI are 
#povvec
#popvec
#order

# AS the inputs to the RCI function simply include a vector giving order by distance, 
# by changing the order vector to density (highest to lowest) a relative densification 
# index (not quite the same as RCO) can be calculated using the same function. 
# The degree of similarity of dissimilarity between RCI and RDI can be taken as an indicator 
# of polycentricity (maybe)

dta %>% 
  inner_join(dist_to_centre) %>%
  mutate(pdens = pop_total / area) %>% 
  group_by(place, year) %>% 
  nest() %>% 
  mutate(
    pvec = map(data, ~ .[["pop_id"]]),
    tvec = map(data, ~ .[["pop_total"]]),
    ordr = map(data, ~ order(.[["distance"]])),
    ordr_dens = map(data, ~ order(.[["pdens"]], decreasing = T))
         ) %>% 
  mutate(rci_val = pmap_dbl(list(pvec, tvec, ordr), RCI)) %>% 
  mutate(rdi_val = pmap_dbl(list(pvec, tvec, ordr_dens), RCI)) %>% 
  mutate(d_val = pmap_dbl(list(pvec, tvec), D)) %>% 
  select(place, year, rci = rci_val, rdi = rdi_val, d = d_val) %>% 
  ungroup -> rci_by_year_place

# Same but for Scotland
dta_scot %>% 
  inner_join(dist_to_centre_scot, by = c("dz_2011" = "dz")) %>%
  mutate(pdens = pop_total / area) %>%
  select(dz_2011, place, year, pop_id = pop_incomedeprived, pop_total, distance, pdens) %>% 
  group_by(place, year) %>% 
  nest() %>% 
  mutate(
    pvec = map(data, ~ .[["pop_id"]]),
    tvec = map(data, ~ .[["pop_total"]]),
    ordr = map(data, ~ order(.[["distance"]])),
    ordr_dens = map(data, ~ order(.[["pdens"]], decreasing = T))
  ) %>% 
  mutate(rci_val = pmap_dbl(list(pvec, tvec, ordr), RCI)) %>% 
  mutate(rdi_val = pmap_dbl(list(pvec, tvec, ordr_dens), RCI)) %>% 
  mutate(d_val = pmap_dbl(list(pvec, tvec), D)) %>% 
  select(place, year, rci = rci_val, rdi = rdi_val, d = d_val) %>% 
  ungroup -> rci_by_year_place_scot

# Combine England w/ Scotland

rci_by_year_place <- bind_rows(
  rci_by_year_place, 
  rci_by_year_place_scot
) %>% 
  filter(place != "Cardiff")
```

## Including Plots

You can also embed plots, for example:

```{r 3d_function, echo=FALSE, cache = T}

show_3d_change <- function(PLACE){
  rci_by_year_place %>% 
    filter(place == PLACE) %>% 
    arrange(year) -> tmp

  n <- nrow(tmp)

  with(tmp, rgl::plot3d(rdi, rci, d, type = "l", expand = 1.10))

  with(tmp, rgl::plot3d(rdi, rci, d, type = "s", 
                        col = c("red", rep("black", n - 2), "blue"), 
                        alpha = c(1, rep(0.5, n-2), 1), add = T))
  rgl::planes3d(1,0,0, 0, alpha = 0.2, add = T)
  rgl::planes3d(0,1,0, 0, alpha = 0.2, add = T)
  title3d(PLACE)
  NULL
}
```

```{r do_glasgow, webgl = T}
show_3d_change("Glasgow")

```
show_3d_change("London")
show_3d_change("Edinburgh")
show_3d_change("Leicester")
show_3d_change("Cambridge")
show_3d_change("Leicester")

### 2017-02-16 scripts/calc_rci_by_place_imd.R

@@ -11,6 +11,7 @@ require(pacman)
 pacman::p_load(
   tidyverse,
   readr, readxl,
+  forcats,
   stringr,  
   purrr,
   rgeos,
@@ -20,6 +21,10 @@ pacman::p_load(
 
 source("scripts/from_gavin/RCI.R")
 
+
+
+
+# Dissimilarity index 
 
 D <- function(minority, total){
   majority      <- total - minority
@@ -35,6 +40,22 @@ D <- function(minority, total){
   out
 }
 
+# Work out the spatial concentration of people in area
+calc_gini <- function(DATA){
+  
+  df <- data_frame(x = c(0, DATA$area), y = c(0, DATA$pop_total)) %>% 
+    arrange(x) %>% 
+    mutate(
+      sx = cumsum(x) / sum(x),
+      sy = cumsum(y) / sum(y)
+    )
+  
+  fn <- approxfun(df$sx, df$sy)
+  auc <- integrate(fn, 0, 1)$value
+  print(auc)
+  out <- auc - 0.5
+  out
+}
 
 
 # imd
@@ -51,11 +72,6 @@ dta_scot <- read_csv("data/simd/simd_combined_on_2011.csv")
 #order
 
 
-# AS the inputs to the RCI function simply include a vector giving order by distance, 
-# by changing the order vector to density (highest to lowest) a relative densification 
-# index (not quite the same as RCO) can be calculated using the same function. 
-# The degree of similarity of dissimilarity between RCI and RDI can be taken as an indicator 
-# of polycentricity (maybe)
 
 
 dta %>% 
@@ -63,7 +79,9 @@ dta %>%
   mutate(pdens = pop_total / area) %>% 
   group_by(place, year) %>% 
   nest() %>% 
+  filter(place != "Cardiff") %>% 
   mutate(
+    pop_area_conc = map_dbl(data, calc_gini),
     pvec = map(data, ~ .[["pop_id"]]),
     tvec = map(data, ~ .[["pop_total"]]),
     ordr = map(data, ~ order(.[["distance"]])),
@@ -72,7 +90,7 @@ dta %>%
   mutate(rci_val = pmap_dbl(list(pvec, tvec, ordr), RCI)) %>% 
   mutate(rdi_val = pmap_dbl(list(pvec, tvec, ordr_dens), RCI)) %>% 
   mutate(d_val = pmap_dbl(list(pvec, tvec), D)) %>% 
[…]
